### 2020-02-28 data-raw/ep1162_prelimDataScreening.Rmd
DBH interpolation

---
title: "EP1162_Prelim_Data_Screening"
author: "Hardy Griesbauer"
date: "18/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)

# Load data
load("../data/ep1162_Data.RData")

```

## Preliminary screening
I did some preliminary screening, to confirm data entry and detect any potential outliers.  Some findings below.

## Outliers
Quick boxplot to confirm DBH and height measurements

#### DBH

```{r echo=FALSE}

dat %>% 
  ggplot()+
  aes(group=Species,y=DBH,fill=Species) +
  geom_boxplot()

```

DBH measurements look reasonable, and no major outliers.  Note that there is a tree with species recorded as "l".

#### Height

```{r echo=FALSE}

dat %>% 
  ggplot()+
  aes(group=Species,y=Height,fill=Species) +
  geom_boxplot()

```
Again, measurements look reasonable here.

### Data entry

One tree has species recorded as "l".  I assume that it's SAF, but might be worthwhile to verify with the field cards?

```{r echo=FALSE}

dat %>% 
  filter(Species=="l") %>% 
  knitr::kable()

```

*ALRF staff confirmed that the species is "Bl"*, so we'll change the species code accordingly in the next line of code:

```{r}
dat<-
  dat %>% 
  mutate(Species=replace(Species,Species=="l","Bl"))

```

There are also two trees recorded with id=631:

```{r echo=FALSE}
# Look at entries where tree number is duplicated within plot
dat %>% 
  add_count(Plot,TreeID,Year) %>% # get a number by this grouping
  filter(n>1) %>% 
  drop_na(TreeID) %>% 
  filter(TreeID!="New") %>% 
    knitr::kable()

```

Looks like it is the same tree (exact same DBH and Height measured in 2019), so assume it's safe to delete the duplicated entry?

*Confirmed that this tree was duplicated by mistake.  We will remove one entry.*

```{r}
dat<-
  dat %>% 
  filter(!(Plot==10 & TreeID==631 & Sector=="NA")) # remove entry

```

There are also quite a few trees with their ID recorded as "New".  Not an issue from an analysis point of view,

### Decreasing DBH 
I also screened trees that had a smaller DBH measurements than a previous measurement at some point.  Below are some examples:

```{r echo=FALSE}
# Look at entries where recorded DBH is less from one measurement to the other

dbhDecrease<-
  dat %>% 
  filter(TreeID!="New") %>% 
  filter(TreeID!=631) %>% # this is a duplicated tree
  pivot_wider(id_cols=c(Plot,TreeID),names_from="Year",values_from="DBH") %>% 
  mutate(diff.2019=`2019`-`2009`) %>% 
  mutate(diff.2009=`2009`-`1997`) %>% 
  mutate(diff.1997=`1997`-`1994`) %>% 
  mutate(diff.1994=`1994`-`1992`) %>% 
  filter(diff.2019<0 | diff.2009<0 | diff.1997<0 | diff.1994<0) %>% 
  dplyr::select(Plot,TreeID) %>% # select these trees with decreasing DBH measurements
  left_join(dat,by=c("Plot","TreeID")) %>% 
  mutate_at(vars(contains("DBH")),round,1) %>% # round to one decimal place
  dplyr::select(Plot:DBH)

  # Show example rows
  dbhDecrease %>% 
  slice(1:20) %>% 
  knitr::kable()

```

In most cases, the differences are very small, and likely can be atrributed to slight measurement differences between years.

In total, there are `r dbhDecrease %>% count(TreeID) %>% nrow()` trees with this situation.  I have the treeID saved and can provide them if we want to double check these trees.

*****
## Trees without monotonically increasing DBH measurements
After discussion, we will handle trees without monotonically increasing DBH measurements as follows:

- Where DBH decreases occur in dead trees or during periods where trees died, we will treat these decreases as likely due to wood shrinkage/rot associated with mortality.
- Where DBH decreases occur during periods where trees were live, we will treat these decreases as a measurement error.  We will use linear interpolation to correct these errors.

Code below.

### Trees with decreasing DBH over dead periods

```{r}
# Step one: identify periods where DBH decrease occured during a period when tree waslive

dbhDecrease %>% 
  filter(Status=="Dead") %>% # filter trees with decreasing DBH that had dead status at any point
  left_join(dat,by=c("Plot","TreeID")) %>%  # join with dataset
  dplyr::select(Plot,TreeID,Year=Year.y,Status=Status.y,DBH=DBH.y)

```

A visual inspection reveals that the following trees had DBH decreases during periods where the tree was dead:

- Plot 3, Tree 457
- Plot 10, Tree 409

These trees should be filtered from further adjustment.

```{r}
# Filter out trees with DBH decrease with dead periods
dbhDecrease<-
  dbhDecrease %>% 
  filter(!(Plot==3 & TreeID==457)) %>% 
  filter(!(Plot==10 & TreeID==409))

```

### Trees with decreasing DBH over live periods
We will make adjustments on a tree by tree basis.

First, reshape the dbhDecrease dataframe to make it easier to see DBH by year, then save it to a .csv file for manual editing.

*I set the following chunk not to run, because we only want to do this once*

```{r eval=FALSE}
dbhEdit1<-
  dbhDecrease%>% 
  dplyr::select(-Status) %>% # Don't need this column anymore
  pivot_wider(names_from="Year",values_from="DBH") %>% 
  write.csv(file="dbh_for_edit.csv",row.names=FALSE)

# Now edit .csv file manually

```

One thing to note: the following trees had fallen as of 2019, which means we should not interpolate their 2019 values:
```{r}
dbhDecrease %>% 
  filter(Year==2019 & is.na(DBH))

```

Also, there are trees that were <7.5cm DBH during the 1992 or 1994 measurements.  DBH values for those trees/periods should not be interpolated either.

For values that shouldn't be interpolated, I entered 'DoNotReplace' in the corresponding cell.

Values to be replaced by interpolation had 'Interpolate' entered into their cell.

```{r}
dbhEdit2<-
  readr::read_csv(here::here("data-raw","dbh_for_edit.csv"))

# save this file so we can go back to it
save(dbhEdit2,file=paste("editedDBH",Sys.Date(),"RData",sep="."))

# Show data with replacements
dbhEdit2

```

Now we can interpolate the missing values

```{r}
library(imputeTS)
library(padr)
library(lubridate)
library(Hmisc)

# Create a vector of years that are not to be interpolated
DoNotReplace<-  
  dbhEdit2 %>% 
  pivot_longer(cols=`1992`:`2019`,names_to="Year",values_to="DBH") %>% 
  filter(DBH=="DoNotReplace")

# Interpolation
dbhEdit3<-
  dbhEdit2 %>% 
  pivot_longer(cols=`1992`:`2019`,names_to="Year",values_to="DBH") %>% 
  mutate(DBH=as.numeric(DBH)) %>% 
  mutate(Year=paste("01-01",Year,sep="-")) %>% 
  mutate(Year=dmy(Year)) %>% 
  group_by(TreeID) %>% 
  pad() %>% 
  ungroup() %>% 
  mutate(Year=year(Year)) %>% 
  as.data.frame()

# Define function to predict DBH values
dbhInterp<-function() {

  YearList=data.frame(Year=c(1992,1994,1997,2009,2019))

 x<-
   dbhEdit3 %>% 
   lm(DBH~Year,data=.) %>% 
   predict(.,newdata=YearList)

 return(data.frame(Year=YearList$Year,DBH.pred=x))

}

# Run function by TreeID
dbhPred<-by(dbhEdit3, dbhEdit3[,"TreeID"],
   function(x) predict(lm(DBH ~ Year, data = x),new=data.frame(Year=c(1992,1994,1997,2009,2019))),simplify=F)

do

```