### 2020-03-20 analysis/heightModel.Rmd
add more files and heigh:dbh differential

---
title: "Height estimates from DBH:Height relationship"
author: "Hardy Griesbauer"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(tidyverse)
library(here)

# Load data
load(here("data","ep1162_Data.RData"))

```

## Estimating tree heights based on dbh:height relationships
In order to estimate stand-level volumes for each treatment, we need to first estimate heights for trees that do not have height measurements.  We will do the following:

1. Model DBH:Height relationships separately for each measurement period, and for Bl:Sx separately (at first).
2. Compare Bl and Sx relationships, compare estimates between years, and determine if we can pool species and/or years to estimate Height.

### Step 1: Explore height:dbh relationships with scatterplots
First, let's explore scatterplots for each species and measurement year:

```{r dbh:height scatterplots,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,fill=Species)+
  geom_point()+
  facet_wrap(Year~Species,ncol=2)+
  theme(legend.position = "none")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

It looks like there are differences between measurement years.  There may be subtle differences between species within a measurement year.  Let's explore that now by superposing Bl and Sx graphs within same year:

```{r dbh:height scatterplots species superposed,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,color=Species)+
  geom_point()+
  facet_wrap(~Year)+
  theme(legend.position = "bottom")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

We can see subtle differences between Bl and Sx DBH:Height relationships in first three measurement periods.  By 2019, it looks like DBH:height relationships converged between the species.

***

Let's look at differences between years by superposing different year graphs for species separately:

```{r dbh:height scatterplots years superposed,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,color=Year)+
  geom_point()+
  facet_wrap(~Species,ncol=2)+
  theme(legend.position = "bottom")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

Looks like Bl maintained the same DBH:Height relationship over all measurement periods.  Sx 2019 relationship differs from other time periods.

****
### Can we pool species and measurement years?

I believe we can test this statistically:

*Note: I will get Peter Ott to confirm this is the correct way to test regression models between groups*

```{r test for species and year effect}

# Assign data
x<-
   dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
    drop_na(Height) # remove trees where height was not measured

# fit a model with all terms
mod.full<-lm(Height~log(DBH)+log(DBH):Species+log(DBH):Year+log(DBH):Species:Year,data=x)

# fit model with only DBH and species terms
mod.species<-lm(Height~log(DBH)+log(DBH):Species,data=x)

# fit model with only DBH and year terms
mod.year<-lm(Height~log(DBH)+log(DBH):Year,data=x)

# fit model with only DBH terms
mod.dbh<-lm(Height~log(DBH),data=x)

# Run through these iteratively
anova(mod.full,mod.dbh)

```

Based on this ANOVA table, there is a significant reduction in residual sum of squares between the full model and the DBH-only model.  The species+DBH and year+DBH models don't seem to add much (not shown).

***
### Compare R^2^ between models
How much variance is explained by each model?  Summary below:

```{r extract r2 and other stats for each model}
rbind(data.frame(Model="full",broom::glance(mod.full)),
      data.frame(Model="dbh-only",broom::glance(mod.dbh)),
      data.frame(Model="dbh+species",broom::glance(mod.species)),
      data.frame(Model="dbh+year",broom::glance(mod.year))

      ) %>% 
  dplyr::select(1:7)

```

Based on this, and the extremely modest increase in R^2^ from DBH-only to full model, I'm tempted to say that we can pool all species and years into the most parsimonious (ie., DBH:Height) model.  But definitely up for discussion!

### DBH:Height equation for pooled data
If we decide to use pooled data for the DBH:Height equation:

```{r dbh:height equation}

paste("Predicted.Height = ",coef(mod.dbh)[2],
      " x ln(DBH) ",
      coef(mod.dbh)[1])
```

Here are the coefficients from Mike's analysis:

*Subalpine fir*
HT = 11.188 x ln(DBH)  - 17.609

*Hybrid white spruce*
HT = 11.79 x ln(DBH)  - 19.637

The coefficients using the pooled species/years are pretty similar to Mike's formula for spruce.