### 2020-03-27 analysis/heightModel.Rmd
volume estimates

---
title: "DBH:Height models for EP1162"
author: "Hardy Griesbauer"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)

# Libraries
library(tidyverse)
library(here)

# Load data
load(here::here("data","ep1162_Data.RData"))

```

## Estimating tree heights based on dbh:height relationships
In order to estimate stand-level volumes for each treatment, we need to first estimate heights for trees that do not have height measurements.  We will do the following:

1. Model DBH:Height relationships separately for each measurement period, and for Bl:Sx separately (at first).
2. Compare Bl and Sx relationships, and determine if we can pool species to estimate Height.

We will generate different dbh:height models for each measurement year, as pooling years would violate assumptions of independence.

### Data preparation: remove trees with broken tops
Trees with broken or dead tops that have height measurements should not be included in DBH:height models.  We can summarise and remove these:

```{r}
dat %>% 
  filter(!is.na(DBT)) %>% # filter for trees with recorded broken top
  filter(!is.na(Height)) %>% # filter for trees with height measurements
  group_by(Species,Year) %>% # summarise trees with broken tops at the species and year level
  summarise(Num.DBT=n()) %>% 
  pivot_wider(names_from="Species",values_from="Num.DBT")

```

### Step 1: Explore height:dbh relationships with scatterplots
First, let's explore scatterplots for each species and measurement year:

```{r dbh:height scatterplots,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% # filter for live trees
    filter(is.na(DBT)) %>% # filter for trees with no broken tops
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,fill=Species)+
  geom_point()+
  facet_wrap(Year~Species,ncol=2)+
  theme(legend.position = "none")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

It looks like there are differences between measurement years.  There may be subtle differences between species within a measurement year.  Let's explore that now by superposing Bl and Sx graphs within same year:

```{r dbh:height scatterplots species superposed,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
   filter(is.na(DBT)) %>% # filter for trees with no broken tops
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,color=Species)+
  geom_point()+
  facet_wrap(~Year)+
  theme(legend.position = "bottom")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

We can see subtle differences between Bl and Sx DBH:Height relationships in first three measurement periods.  By 2019, it looks like DBH:height relationships converged between the species.

***

Let's look at differences between years by superposing different year graphs for species separately:

```{r dbh:height scatterplots years superposed,fig.height=12, fig.width=10}

  dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
   filter(is.na(DBT)) %>% # filter for trees with no broken tops
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,color=Year)+
  geom_point()+
  facet_wrap(~Species,ncol=2)+
  theme(legend.position = "bottom")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

Looks like Bl maintained the same DBH:Height relationship over all measurement periods.  Sx 2019 relationship differs from other time periods.

As per previous discussion, we shouldn't pool measurements across measurement years because this would violate assumption of independence among samples.

****
### Can we pool species?

I believe we can test this statistically:

*Note: I will get Peter Ott to confirm this is the correct way to test regression models between groups*

```{r test for species effect}

# function to compare three models
# separate models will be fit for each year

speciesMod1<-function(year) {

# Assign data
x<-
   dat %>% 
    filter(Species %in% c("Bl","Sx")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% 
   filter(is.na(DBT)) %>% # filter for trees with no broken tops
  filter(Year==year) %>% 
    drop_na(Height) # remove trees where height was not measured

# fit a model with all terms
mod.full<-lm(Height~log(DBH)+log(DBH):Species,data=x)

# fit model with only DBH terms
mod.dbh<-lm(Height~log(DBH),data=x)

# extract coef and other values
out<-data.frame(year=year,intercept=coef(mod.dbh)[1],
                slope=coef(mod.dbh)[2],
                p.value=broom::glance(mod.dbh)$p.value,
                r2=broom::glance(mod.dbh)[1])
rownames(out)=year

# Return
return(list(anov=anova(mod.full,mod.dbh),out=out))

}

```

Now let's run the model for the four measurement periods:

1992:
```{r}
speciesMod1(1992)$anov
```

***
1994:

```{r}
speciesMod1(1994)$anov
```

***
1997:

```{r}
speciesMod1(1997)$anov
```

***
2019:

```{r}
speciesMod1(2019)$anov
```

For each of the four measurement periods, there is no significant difference between models that include DBH:species interactions and just use DBH alone.  Therefore, we can develop a single height model for each measurement period.

***

### DBH:Height equation for pooled species data

```{r dbh:height equation}

heightModels<- # assign 
  rbind(speciesMod1(1992)$out,
      speciesMod1(1994)$out,
      speciesMod1(1997)$out,
      speciesMod1(2019)$out) %>% 
  as_tibble()

print(heightModels)

```

#### Height model for 2009 measurement period
Height was not recorded during the 2009 measurements, however, we should be able to estimate heights using the 2009 DBH data.  I propose the following model to estimate 2009 heights:

### Graph out intercept and slope for four measurement period
I noticed that the intercept and slope vary monotonically with years, which suggests that we might be able to predict intercept/slope for a given year (in this case 2009).  Let's examine this with a scatterplot:

```{r}

heightModels %>% 
  dplyr::select(1:3) %>% #select first three columns
  pivot_longer(-year,values_to="coefficient",names_to="component") %>% # pivot longer for facet plot
  ggplot()+
  aes(x=year,y=coefficient)+
  geom_point()+
  geom_line(color="red")+
  geom_vline(xintercept=2009,linetype="dashed")+
  facet_wrap(~component,nrow=2,scales="free")

```

Based on these relationships, it seems we can reasonably estimate the slope and intercept for a dbh:height model at year 2009.  Thoughts?

### Calculate 2009 intercept and slope coefficients for dbh:height model

```{r}

# Calculate slope to estimate intercept:
int.slope<-
  with(heightModels,heightModels[year==2019,"intercept"]-heightModels[year==1997,"intercept"])/(2019-1997)
  int.slope$run=diff(c(1997,2009))
  int.2009<-  as.numeric(with(int.slope,run*intercept)+with(heightModels,heightModels[year==1997,"intercept"]))

####
slope<-
  with(heightModels,heightModels[year==2019,"slope"]-heightModels[year==1997,"slope"])/(2019-1997)
  slope$run=diff(c(1997,2009))
  slope.2009<-  as.numeric(with(slope,run*slope)+with(heightModels,heightModels[year==1997,"slope"]))

### Add these values into heightModels table
  heightModels<-
    heightModels %>% 
    add_row(year=2009,intercept=int.2009,slope=slope.2009) %>% # add 2009 coefficients to table
    arrange(year) # order by year

# print to screen
  heightModels

```

### Save file for volume estimates
```{r}
# save coefficients for volume estimates
save(heightModels,file=here("data","heightModels.RData"))
```