### 2021-05-06 manuscript_draft.Rmd
updtae negExp

---
title: "Manuscript_draft"
author: "Michael Jull and Hardy Griesbauer"
date: "10/12/2020"
output:
  word_document:
    reference_docx: wordTemplate2.docx
  html_document:
    df_print: paged
csl: cjfr.csl
bibliography: C:\\Users\\hgriesba\\Google Drive\\Papers\\library.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r load libraries for analyses, echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE,ft.align="left")

# Load libraries
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom)
library(MuMIn)
library(ggeffects)
library(car)
library(flextable)
library(captioner)
library(lsmeans)
library(cowplot)
library(officer)
library(emmeans)

load("data/ep1162_Data.RData")
load("data/ep1162_plotData.RData")

# SOME PRELIM DATA FORMATTING
# formatting plotDat
plotDat<-
  plotDat %>% 
  mutate(Plot=as.factor(Plot))

# set up diameter classes
diamClass=seq(from=7.5,to=100,by=5)

# Set against scientific notation
options(scipen=999)

# Merge data
X<-
  dat %>% 
  left_join(plotDat,by="Plot") %>% 
  mutate(DBH.class=cut(DBH,diamClass,include.lowest=TRUE,ordered_result = TRUE,right=F,
                       labels=diamClass[1:(length(diamClass)-1)])) %>% 
  mutate(Plot=as.factor(Plot)) %>%

  # replace species codes with words
  mutate(Species=recode_factor(Species,Bl="Fir",Sx="Spruce")) %>% 
  mutate(Species=fct_drop(Species)) %>% # have an extra level in here for some reason

  # Create TU lookup
  mutate(BA.Target=paste0("TU:",BA.Target)) %>%

  # need to remove plots 16-19 because they were only measured once
  filter(!Plot %in% c(16,17,18,19))

# set up captions

# Code to create captions for tables
tableNums <- captioner(prefix = "Table")

# Code to create captions for figures
figNums <- captioner(prefix="Figure")

# Code to create captions for appendices
appNums <- captioner(prefix="Appendix")

# Set ggplot theme
theme_set(theme_bw())

# Define function to convert diameter (cm) to basal area (m2), and vice versa
d2ba<-function(diam) {return((diam/2/100)^2*pi)}
ba2d<-function(ba) {return(sqrt(ba/pi)*200)}

# plot multiplier
pm=20

```

# To-do list

1. COmplete Q factor coding

### Comments from Mike's emails

we can put text below each comment to describe how it was addressed, and
then delete comments for any issues once they are resolved.

#### Use of terminology

-   The term "units" is used in several places, where it seems that you
    mean "treatments" (or treatment types, treatment unit).

-   The use of the term "plot" or "plots" is used appropriately as far
    as I can see.

-   The term "fir" is used to describe subalpine fir or 'balsam' in
    several parts of the text, including possibly figure and table
    captions. I know what species you are referring to, of course, but
    others, especially cursory readers, may misconstrue it. The
    colloquial use of 'fir" is spread across both Abies and Douglas-fir.
    "Balsam" as a term has well-documented ambiguities, and while
    Canadians tend to use balsam for Abies, Americans often call Abies
    "firs", as in the various species of true fir. I would recommend
    using Abies as the shorthand descriptor in the text and captions for
    subalpine fir. It is a short word, and we have only 1 Abies species
    at this site. Also, my experience in the on-line-search literature
    world and with things like Google Scholar is that Abies is a much
    better and more diagnostic search word for this subject matter than
    fir or balsam. "Spruce-Abies" is very unambiguous as a descriptor
    for this forest type.

-   Rate of change (increase, decrease) in vol, BA, and sph: In both in
    your stats tables and figure captions, and in the text narrative,
    ensure that what you are describing is clearly and accurately
    described as either: (a) the incremental change in values between
    two measurement periods, or (b) the difference in absolute value in
    certain parameters (e.g. standing volume) between two measurement
    periods, or between two treatments. Also, the annualized growth rate
    of something averaged over a measurement period (e.g. mean PBAI,
    like BA growth in m2/yr over a period) is different than the rate of
    change in a growth rate over time (e.g. is a growth rate increasing
    or decreasing). I got confused a few times on this issue.

### Commentary on sections and sub-sections

#### Section 3.1.1 BA increment from 1992 to 2019

-   Great: Very clear and straightforward. i.e. - log-transformed
    treatment (initial BA) was a very statistically significant
    predictor of tree response across dbh classes for both Sx and Bl.
-   Don't know what all the random effects outputs are. I am sure they
    are meaningful, but I just dont know what they mean.
-   Number of observations of 510: Should ensure it is clear that only
    trees \> 7.5 cm dbh are counted.
-   Would be good to define the dbh class limits somewhere. In Appendix
    6 table, we describe the class mid-point.
-   For Appendix 6, I assume that the null hypothesis for the ratio is
    = 1.

#### Section 3.1.2 Mortality

-   I am thinking that mortality should / could be analysed as a
    frequency response - i.e. - expected vs actual distribution of
    mortality by size class and/or species. e.g. a Chi-squared or allied
    type of analysis. Null hypothesis is that mortality of stems is
    randomly distributed across different sizes and species of trees. H1
    is that it is non-random.
-   I'd suggest that we don't compare total mortality (e.g \# of stems)
    across treatment units, but rather frequency relative to initial
    number of stems (i.e. - at beginning of observation period). That
    way we normalize mortality rates relative to initial stem densities.
    Otherwise, results could be mis-leading - e.g. Control treatments
    could have more mortality in overall sph, simply because they have
    more initial stems.
-   ACTION ITEM FOR ME: I need to check my original datasets and data
    sorting iterations to see how I handled mortality observations for
    the 1992 and 1994 plot remeasurements.
-   Note for you: All of the mortality you've analysed and identified so
    far is mortality that is incremental following 92/94, so it is still
    solid analysis data, just with that qualification. I will clarify
    the 92/94 mortalilty observations for the Methods section; we did
    observe dead trees at these times, but these mortality pre-dated our
    treatments and/or first measurements. Anyway, probably not a big
    deal, but this needs examination and clarification to explain this.
    If necessary, I will quantify observations of dead trees observed on
    these years, because they are actually \> 0.

Testing for density-dependence of mortality responses... some ideas or
hypotheses.. Note that f (\_\_\_) means "a function of"...

-   MORT RATE of a size class = f (Sum of BA in that plot \> than that
    size class). Would assume asymmetric competition where only bigger
    trees exert competition on smaller trees.
-   MORT RATE of a size class = f ( Sum of BA in that plot that is
    either \> or \< that size class). Assumes symmetric competition
    where both big and small trees exert competition on their neighbors.

Caveats: (a) our plots and tree distributions might be too spatially
variable to get a reliable test of these hypotheses, and (b) BA would be
the BA at the beginning of a measurement period. And (c) a tree that
died of suppression would probably show reduced growth rates prior to
death; but trees that had abiotic damage and mortality - e.g. - snow or
wind breakage - might not. Complicated stuff !!

#### 3.2.1 Estimating tree height from diameter

-   Really solid stats and rationale for a single ht:dbh model. Great !
-   What is really interesting and useful about this is that the ht:dbh
    relationship is seemingly statistically independent of treatment or,
    that is, initial density, A very useful simplifying assumption
    proven for future modellers of these stands. We should highlight
    this in either our results or discussion.

#### 3.2.2 Plot means

-   See notes above about use of term "units" vs treatment (or similar).
-   Clarify effects of treatment and/or species on growth rate vs
    standing vol / BA at each measurement period.
-   A Figure 7b or equivalent could visually present changes in growth
    rate over time. (e.g. previous graphs we've developed)
-   Under Observations, \# 2 - the 2nd sentence needs work to edit and
    clarify.

#### 3.2.3 Stand attributes by diameter class

-   We should define what we mean by the term inverse J or inverse
    J-shaped dbh distribution. Just for clarity.
-   Should we quantitatively describe these distributions? Don't have
    to, just asking the question.
-   Is the merch limit for min merch tree size 18.5 cm dbh, or 17.5 cm
    dbh? I remember the latter from various cruises and cruise
    compilations.
-   There are two Figure 2's, one is for vol and one is for sph.
-   Clarify in figures and text that sph is based on trees 7.5 cm dbh
    and greater. Increases in sph can be from ingress of trees growing
    from size classes below this limit.
-   A suggestion: Have a table comparing total volume to merch volume by
    treatment, etc? Would be useful for some readers, and communicate
    MAI and volume growth rates too.

#### 3.2.4 Species composition

-   First sentence: Clarify what you mean by "organized similarly". Does
    this refer to frequency distribution ? What constitutes similarity?

#### Mortality

Last winter, as well as keeping reference copies of the initial "raw"
dataset following data entry and related stuff, I kept copies each stage
of data cleaning and prep (about 10 iterations).

We did have "dead tree" observations in the first \~ 2 year of
monitoring. The key thing to see is : if I can determine is when the
trees were dead, and/or when death was initiated, and why.\
Timing of tree death and observations of causal factors of tree death in
these first two years will help inform our interpretation and data
handling of these observations.

If trees were dead or moribund at the first "initial" observation or two
(as soon after the logging as possible), this is -- arguably - different
than mortality that occurs over time in response to the basal area
treatment and/or could be considered to be density- or
treatment-dependent in some way.

The 1992 data was in mid-March really early after logging, and before
snowmelt (in retrospect not a great idea but we had resources available
before the end of the fiscal year at the time). The 1994 data was from
when the plots were re-visited under snow-free conditions. We did a
thinning and sanitation culling of trees in the treatment units and
PSP's in 1994 to remove logging-damaged trees and most deciduous, to
bring the units and PSP's down to their final treatment BA. 1994
remeasurements were completed after that.

A few plots, as you know, were established in 1994 not 1992.

Interpretive implications:

1.  The very-early observed mortality may be a result of pre-existing
    stand conditions not the created, treatment-induced stand
    conditions. "stochastic" vs potentially treatment induced.
2.  It can be argued that trees dead at time of treatment or very soon
    thereafter were never taking up growing space in the post-harvest
    stand, and therefore would not contribute to the initial "green"
    basal area going forward.
3.  We do have 1992 data in most plots, but it was 1994 when the BA in
    each trmt unit and embedded PSP was finalized. 1994 is the most
    homogeneous start point for all PSP's.

Regarding tree mortality in the period from 1992 to 1994 at Summit Lake
EP 1162:

I reviewed our data regarding this, any relevant project records, and my
own recollection of our monitoring of tree mortality during this period.
Here is my train of thought, and what I have concluded on this topic:

1992 Observations and PSP establishment limitations

Trees that were dead at the time of the 1992 PSP establishment were not
tagged or recorded. The rationale for this was that these trees were
dead and therefore not taking up up growing space in the stands. Dead
trees in 1992 would have been either snags / dead trees pre-existing the
harvest, or trees fatally damaged or killed by logging activities (e.g.
-- trees with stems snapped and crowns broken off, no live limbs) or
trees smashed to the ground by felled trees).

Summer 1994 sanitation treatment and Fall 1994 PSP Remeasurement

EP 1162 treatment units, including the PSP area contained therein, were
assessed in Spring 1994 to review stand density / basal area, and stand
condition, in order to:\
a) identify and mark-to-cut residual trees that were considered too
badly scarred or damaged to constitute good growing stock for trial
purposes, b) where needed, to lower the initial BA closer to the target
level, c) to either fell or girdle birch trees, especially in the PSP's
themselves, but usually also within the TU as a whole, and d) to fell
any dead, dangerous, and/or moribund trees within the treatment units
and PSP's. Generally most if not all of these trees felled in the
sanitation felling were understory or subdominant trees, and
occasionally a lower codominant tree. Tagged trees felled by the
sanitation felling were documented on field sheets and our Excel
datasets, so we can retrieve the species and size of these trees as
needed. However, sanitation-felled trees included both live, moribund,
and dead trees -- generally all poorer quality or vigor though.

After felling, the felled trees were bucked to a 2 to 3 m length so as
to lay these felled trees to as close to the ground as possibly,
maximize decay, and minimize physical obstructions within the TU's and
PSP's.

As a result of the sanitation felling, potentially some trees which died
or were damaged between 1992 and Summer 1994 were felled without
recording these mortality events and precise tree vigor beforehand. Some
minor mortality occurred between 1992 and 94, but was obscured by the
post-harvest sanitation treatment.

The 1994 Remeasurement occurred in Fall 1994, with a few plot
remeasurements occurring in Winter 1994/95 (i.e. no intervening growing
season).

For scientific rigour, we should not state that there was zero mortality
between 1992 and 1994. Rather, I suggest, it is more accurate to state
that data on tree mortalilty between 1992 and 1994 was not collected,
and that the Summer 1994 sanitation felling treatment did not allow
accurate documentation of tree mortality in this initial period on a
retrospective basis.

We don't, therefore, have data on mortality between 1992 and 1994,
though my recollection is that it was minor in nature.

The difference in basal area between 1992 and 1994 is predominantly a
product of the sanitation felling of some basal area, with minor natural
mortality.

Mortality occurring between Fall 1994 and Fall 1997 remeasurements was
captured accurately by these two sequential remeasurements.

"The difference in basal area between 1992 and 1994 is a result of basal
area reductions through the sanitation felling of some stems and some
minor natural mortality, combined with basal area increases through the
growth of remaining live stems."

A review of the various plots indicates that between 1992 and 1994, some
PSP's decreased in BA (as sanitation removals and minor mortality
exceeded growth of live stems), while others increased in BA as tree
growth exceeded BA removals.

-   Is there any change in which sizes or size classes of trees die as
    stand volume increases? - e.g. is competition asymmetrical or
    symmetrical? Currently the y axis is total number of trees, I
    believe.
-   Not sure if we have enough sample size of trees to look at species
    dynamics (Sx vs Bl).
-   I am wondering if the same relationship that appears between volume
    and tree mortality, is mirrored in the relationship between BA and
    tree mortality. A bit more on this below...

Foresters (like me) do like to use BA per ha as a density measure for
field planning and stand assessment because it seems to be broadly
useful and is easily assessed via stand walkthroughs or plots. Sapwood
BA is pretty well correlated with foliar biomass of the stand - a
biological linkage. Tree BA is often correlated well with things like
crown width.

I hadn't thought alot previously about vol per ha as a density
measurement, but it does have the characteristic of integrating not just
BA, but also tree height, so it integrates tree growing space occupancy
in both in horizontal dimensions (crown width and probably root spread)
but also vertical. Vol ha theoretically provides more of a 3D measure of
how the trees are occupying a given land area. The fact that volume is
of commercial interest is a bonus. Key assumptions here: That trees are
healthy and most importantly, crowns are health and growing relatively
optimally relative to their growing space opportunities. Once crown
vigor etc starts to decline with age, pests or pathogens, then the
relationship could start to break down.

\newpage

# Introduction

# Methods

This section will describe the methods we used for data analyses, not
field data collection.

From our technical report submitted to FCI earlier:

Our data analyses and summaries examined the effects of the different
levels of stand-level basal area density treatments on three main
attributes of stand development following treatment. These include, in
order:

1.  Post-treatment stand basal area dynamics, including the rate and
    pattern of basal area re-growth and recovery over time following
    different levels of initial basal area density reductions;
2.  Stand structural outcomes, including treatment influences on the
    overall abundance of trees (in sph) by diameter class, within the
    different treatment types, and;
3.  Tree species composition outcomes, including treatment influences on
    the relative abundance and size class distributions of subalpine fir
    and hybrid white spruce within the different treatment types.

Analyses are grouped into three main categories (below). These are
treated as separate categories in the Results section.

1.  Tree-level responses to treatment;
2.  Stand-level responses to treatment:

## Data analyses

### Tree-level growth responses

To test the hypothesis that tree-level radial growth response over the
1994-2019 period varied with interactions between (i) tree diameter at
the time of final treatment (1994); (ii) treatment; and (iii) species,
we fit linear mixed effects models with log-transformed 1994-2019
diameter increment as the response variable, and treatment, species and
log-transformed 1994 diameter as interacting fixed effects. To account
for plot-level productivity differences, we included plot as a random
intercept effect. Models were fit using the 'lme4' package [@Bates2015] for the R statistical environment [@R-base].
We also fit a similar model with the 1994-2019 diameter increment
expressed as a percentage of 1994 diameter. Plots of residuals and fit
lines with orginal data were examined to evaluate model goodness of fit.
Variance explained for each model are reported using the approach
described in Nakagawa et al. [-@Nakagawa2013], as implemented in the
MuMIn package for R [@barton2009mumin]. Pairwise contrasts were
performed using the 'emmeans' package for R [@Lenth2019] and Tukey Honestly Significant Difference test of
*p* value adjustment was used to control for multiplicity. We report all
significant results at p<0.05.

We calculated growth dominance statistic (GDS) for each plot and time period.  Growth dominance in a stand considers the contribution of individual trees to total stand growth as a function of their size [@Binkley2006; @Forrester2019; @West2014].  The GDS expresses growth dominance as a single metric, where a value of 0 indicates a stand where the growth of individual trees, expressed as a proportion of stand total growth (in this case, basal area increment), is equal to their size, expressed as a proportion of stand total size (in this case, basal area).  A negative (positive) GDS indicates a stand where smaller (larger) trees contribute more to stand growth relative to their proportional size.  For more information on the formula used to compute GDS, we refer the reader to West [-@West2014].

### Stand-level growth and structure

#### Stand-level volume and basal area estimates

Total and merchantable volumes were estimated for trees using species-
and region-specific equations [@nigh2016] based on tree height and DBH.
Consistent with forest management practices in British Columbia, we
estimated volumes for trees meeting a minimum diameter utilization limit
of 18.5cm DBH. For trees lacking a tree height measurement, we estimated
height by fitting a single diameter-height model developed with a linear
mixed effects model, with log-transformed DBH as a fixed effect, and
tree nested within plot as random intercepts. Trees with broken tops
were omitted from diameter-height models. We summed tree-level estimates
of volume and basal area to generate stand-level estimates for each
plot. To test the hypothesis that changes in stand attributes over 27
years varied by treatment, we fit separate linear models with stand
total volume, basal area, and density as response variables, and the
interaction between treatment and year as the independent variable.

#### Growth dominance coefficients
*Need to describe GDC methods here*

#### Diameter distributions
We quantified the stand structure in each plot and time period by computing Q factor, which is the ratio between adjacent diameter classes.  Diameter classes were 5cm wide.

# Results

## Tree-level

### Basal area increment from 1994-2019

```{r mixed effects model of tree growth}

# Filter only for trees that have measurements in 1994 and 2019 
treeList<- # assign variable
  X %>% 
  filter(TreeID!="New") %>% # remove new trees
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(Species %in% c("Spruce","Fir")) %>% # filter birch and fdi out of dataset
  filter(Status=="Live") %>% # filter for live trees
  filter(Year %in% c(1994,2019)) %>% #filter for start and end dates
  group_by(TreeID) %>% 
  summarise(N=n()) %>% 
  filter(N==2) %>% #filter for trees with 5 measurements
  pull(TreeID)

# derive periodic increment (period diameter or basal area increment)
piDat<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% treeList) %>% 
  mutate(BA=d2ba(DBH)) %>% # convert DBH to basal area
  dplyr::select(Plot,TreeID,Species,Year,BA) %>% 
  pivot_wider(names_from=Year,values_from=BA) %>% 
  mutate(PI=`2019`-`1994`) %>% # 
  filter(PI>0) %>% # one tree has a PI of 0, has to be removed
  mutate(PIpct=PI/`1994`*100) %>% # express periodic increment a a percentage
  dplyr::select(Plot,TreeID,Species,PBAI=PI,PBAIpct=PIpct,Init=`1994`) %>% 
  left_join(plotDat,by="Plot")

# fit lmer model
  # fixed effects model with species pooled, using PBAI
  fitBA<-
    piDat %>% 
    lmer(log(PBAI)~BA.Target:log(Init):Species+(1|Plot),data=.)

  # create a reference grid for five different diameter classes (converted to BA)
  rg<-ref_grid(fitBA,at=list(Init=d2ba(c(10,20,30,40,50))),type="response",nesting=NULL)

  # Using PBAIpct
  fitBApct<-
    piDat %>% 
    lmer(log(PBAIpct)~BA.Target:log(Init):Species+(1|Plot),data=.)

# create a reference grid for five different diameter classes (converted to BA)
  rg.pbaipct<-ref_grid(fitBApct,at=list(Init=d2ba(c(10,20,30,40,50))),type="response",nesting=NULL)

```

```{r tree level species contrasts}

sppContrast<-

  pairs(rg,by=c("BA.Target","Init"),type="response") %>% 
  as.data.frame() %>%
    mutate(Init=as.numeric(as.character(Init))) %>% # change from factor to numeric
    mutate(Init=ba2d(Init)) %>% # convert basal area to diameter
  arrange(BA.Target,Init) %>% 
  rename(DBH.1994="Init")

```

```{r treatment contrasts}

trtContrast<-
  pairs(rg,by=c("Species","Init"),type="response") %>% 
  as.data.frame() %>% 
  mutate(Init=as.numeric(as.character(Init))) %>% # change from factor to numeric
  mutate(Init=ba2d(Init)) %>% # convert basal area to diameter
  arrange(Species,contrast,Init) %>% 
  rename(DBH.1994="Init")

```

The tree-level basal area increment model indicated that interactions
between log-transformed tree size, species and treatment explained
`r round(r.squaredGLMM(fitBA)[2]*100,1)`% of the variation in
log-transformed 1994-2019 tree-level BAI, and
`r round(r.squaredGLMM(fitBApct)[2]*100,1)`% of the variation in basal
area increment as a percentage of tree size in 1994 (p\<0.0001 for both,
`r tableNums("fitBA.table",display="cite")`). Estimated marginal means
from the model showed that between species and among treatments, basal
area increment increased with tree size, meaning that larger trees
increased basal area more than smaller trees
(`r figNums("treeGrowth.plot",display="cite")`), although smaller trees
increased their size proportionally more than large trees
(`r figNums("treeGrowthPct.plot",display="cite")`). Species contrasts
showed that spruce trees increased basal area significantly more than
than fir in the high-removal and control treatment units, whereas basal
growth between the two species was close to equal in the low-removal
unit (`r appNums("sppContrast.table",display="cite")`). Within-species
contrasts showed that spruce basal area increment in the high-removal
unit exceeded spruce growth in both other units
(`r appNums("trtContrast.table",display="cite")`), whereas fir basal
area increment in the high- and low-removal units was higher than the
control unit, and there were no significant differences between the two
harvested units (`r appNums("trtContrast.table",display="cite")`).

## Stand-level analyses

```{r height to diameter models}

# Fit model
htDiamFit<- 
  X %>% 
    filter(Species %in% c("Fir","Spruce")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% # filter for live trees
    filter(is.na(DBT)) %>% # filter for trees with no broken tops
    drop_na(Height) %>% # remove trees where height was not measured
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create tree ID column
  lmer(log(Height)~log(DBH)+(1|Plot/TreeID),data=.)

# Apply fit model to data and create new height column
X<-
  X %>% 
  mutate(Height.Pred=coalesce(Height,exp(fixef(htDiamFit)[1]+fixef(htDiamFit)[2]*log(DBH))))

```

### Tree height modeling

Tree heights had a positive nonlinear relationship with diameter, for
both species and across all time periods
(`r appNums("heightDiam.plot",display="cite")`). We originally developed
separate models for each species and time period, however, a comparison
of height-diameter models showed that intercepts and slopes did not vary
significantly (i.e., p\>0.05) between species or time periods (not
shown). Therefore, we developed a single height-diameter linear
mixed-effects model using
`r X %>% filter(Species %in% c("Fir","Spruce")) %>% filter(Status=="Live") %>% filter(is.na(DBT)) %>% drop_na(Height) %>% nrow()`
tree height-diameter observations in the dataset. This model explained
`r round(MuMIn::r.squaredGLMM(htDiamFit)[1]*100,1)`% of variation in
log-transformed heights, with an intercept of
`r round(fixef(htDiamFit)[1],3)` and slope coefficient of
`r round(fixef(htDiamFit)[2],3)` \* log-transformed DBH (not shown).
Residual and predicted vs actual height plots were assessed to ensure
goodness of fit (`r appNums("htModelFit.plot",display="cite")`). Using
this model, we generated height estimates for trees without height
measurements, and then estimated tree total volume as per Nigh
[-@nigh2016].

```{r hardcode volume coefficients per Nigh}

volCoef<-

  ## TOTAL VOLUME
  #  Subalpine fir 
  data.frame(Species="Fir",VolType="total",b0=-9.79,b1=1.813,b2=1.033) %>%

  # Douglas-fir
  add_row(Species="Fd",VolType="total",b0=-10.19,b1=1.73,b2=1.206) %>%

  # Add spruce
  add_row(Species="Spruce",VolType="total",b0=-10.024,b1= 1.801,b2= 1.107) %>%

  ## MERCHANTABLE VOLUME
  # Subalpine fir
  add_row(Species="Fir",VolType="merch",b0=-9.980,b1= 1.792,b2= 1.102) %>%

  # Douglas-fir
  add_row(Species="Fd",VolType="merch",b0=-10.503,b1= 1.724,b2= 1.296) %>%

  # Spruce
  add_row(Species="Spruce",VolType="merch",b0=-10.280,b1= 1.769,b2= 1.206 ) %>%

  mutate(BGC.Zone="SBS") %>%

  # convert species to factor
  mutate(Species=factor(Species)) %>% 
  dplyr::select(Species,BGC.Zone,everything())

```

```{r summarize stand attributes by diameter class}

# We will produce diameter class stand attributes for two time periods 1994 and 2019.

standDiam.X<-

  # Determine which plots have measurements in 1994 and 2019
  X %>% 
   filter(Year %in% c(1994,2019)) %>% 
   filter(Status=="Live") %>%

  group_by(Plot,Year) %>% 
  summarise(n=n()) %>% 
  ungroup() %>% 
  group_by(Plot) %>% 
  summarise(n=length(Year)) %>% 
  filter(n>1) %>% # pick plots with measurements in both years
  pull(Plot) %>% 
  as.data.frame() %>% 
  setNames("Plot") %>%

  # Now filter original dataset 
  left_join(X,by="Plot") %>% 
  mutate(Plot=fct_drop(Plot)) %>%  # drop unused levels

  # Filter for years of interest
  filter(Year %in% c(1994,2019)) %>% 
  mutate(Year=fct_drop(Year)) %>% 
  filter(Status=="Live") %>%

    # Calculate volume
  left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  mutate(BA=d2ba(DBH)) %>% 
  dplyr::select(BA.Target,Plot:Species,Year,DBH,BA,vol,DBH.class,Height.Pred) %>% 
  mutate(Species=factor(Species)) %>%

 # Now create plot-level stats
  ungroup() %>% 
  group_by(BA.Target,Plot,Species,Year,DBH.class) %>% 
  summarise(BA=sum(BA)*pm,
            Vol=sum(vol)*pm,
            SPH=n()*pm) %>% 
  ungroup()

# Now summarize periodic diameter class stand attributes by species

# # Create total summaries
#   diamClass.total<-
#     standDiam.X %>% 
#     
#     # create plot totals first
#     group_by(BA.Target,Plot,Year,DBH.class) %>% 
#     summarise(BA=sum(BA),
#               Vol=sum(Vol),
#               SPH=sum(SPH)) %>% 
#     mutate(Species="Total") %>% 
#     ungroup() %>% 
#     
#     # now we have to complete the data frame to create BA.Target means
#     complete(nesting(BA.Target,Plot),Year,DBH.class,fill=list(BA=0,SPH=0,Vol=0)) %>% 
#     group_by(BA.Target,Year,DBH.class) %>% 
# 
#     # Plot means
#     summarise(BA=mean(BA),
#               Vol=mean(Vol),
#               SPH=mean(SPH)) %>% 
#     ungroup() %>% 
#     mutate(Species="Total") %>% 
#     filter(BA>0) %>% 
#     dplyr::select(BA.Target,Species,Year,DBH.class,BA,Vol,SPH)

  diamClass.spp<-
     standDiam.X %>% 
    # filter(Species %in% c("Fir","Spruce")) %>%  
    # create plot totals first
    filter(Species!="Ep") %>% # remove birch to clarify
    mutate(Species=fct_drop(Species)) %>% # drop Ep from levels

    # now we have to complete the data frame to create BA.Target means
    complete(nesting(BA.Target,Plot),Species,Year,DBH.class,fill=list(BA=0,SPH=0,Vol=0)) %>% 
    group_by(BA.Target,Species,Year,DBH.class) %>%

    # Plot means
    summarise(BA=mean(BA),
              Vol=mean(Vol),
              SPH=mean(SPH)) %>% 
    ungroup() %>% 
    filter(BA>0) %>% 
    dplyr::select(BA.Target,Species,Year,DBH.class,BA,Vol,SPH) %>%

    # reorder Species levels for plotting
    mutate(Species=fct_relevel(Species, "Fd", after = Inf)) %>% 
    mutate(Species=fct_recode(Species,`Douglas-fir`="Fd"))

```

### Plot means

```{r generate stand level plot means}

plotData.X<-

  X %>%

  filter(Status=="Live") %>%

    # Calculate volume
  left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  mutate(BA=d2ba(DBH)) %>% 
  dplyr::select(BA.Target,Plot:Species,Year,DBH,BA,vol,DBH.class,Height.Pred) %>% 
  mutate(Species=factor(Species)) %>%

  # Remove volumes below 18.5cm DBH
  mutate(vol=replace(vol,DBH<18.5,0)) %>%

  # now create plot sums grid so that you can calculate means
  group_by(BA.Target,Plot,Year) %>% 
    summarise(Vol.sum=sum(vol,na.rm=T)*pm,
              BA.sum=sum(BA)*pm,
              SPH.sum=n()*pm) %>%

  ungroup() %>%

  # Create QMD
  mutate(QMD.mean=sqrt(BA.sum/(SPH.sum*0.0000785)))

# Now summarise plot data at the treatment unit level
standData.X<-
  plotData.X %>% 
  group_by(BA.Target,Year) %>% 
  summarise(Vol=mean(Vol.sum),
            Vol.se=sd(Vol.sum)/sqrt(length(Vol.sum)),
            BA=mean(BA.sum),
            BA.se=sd(BA.sum)/sqrt(length(BA.sum)),
            SPH=mean(SPH.sum),
            SPH.se=sd(SPH.sum)/sqrt(length(SPH.sum)),
            QMD=mean(QMD.mean),
            QMD.se=sd(QMD.mean)/sqrt(length(QMD.mean)))

```

```{r periodic increment by treatment level}

# Create a function to compute standard error for vectors with NA
incSE<-function(x) {

  x2<-
    x %>% 
    .[!is.na(.)]

    sd(x2)/sqrt(length(x2)) %>% 
    return()
  }

perInc.X<-
  plotData.X %>% 
  pivot_longer(cols=Vol.sum:QMD.mean,names_to="var",values_to = "val") %>% 
  pivot_wider(names_from = Year,values_from = val) %>% 
  mutate(Period1=(`1997`-`1994`)/3,
         Period2=(`2009`-`1997`)/11,
         Period3=(`2019`-`2009`)/11) %>% 
  dplyr::select(BA.Target:var,Period1:Period3) %>% 
  pivot_longer(cols=Period1:Period3,names_to="Period",values_to="Values") %>% 
  pivot_wider(names_from=var,values_from=Values) %>%

  # now summarize at treatment unit level
  group_by(BA.Target,Period) %>% 
   summarise(Vol=mean(Vol.sum,na.rm=T),
            Vol.se=incSE(Vol.sum),
            BA=mean(BA.sum,na.rm=T),
            BA.se=incSE(BA.sum),
            SPH=mean(SPH.sum,na.rm=T),
            SPH.se=incSE(SPH.sum),
            QMD=mean(QMD.mean,na.rm=T),
            QMD.se=incSE(QMD.mean)) %>%

  # rename periods
  mutate(Period=dplyr::recode(Period,
                       Period1="1994-1997",
                       Period2="1997-2009",
                       Period3="2009-2019"))

```

```{r fit model to plot data}

# Rename columns in plotData.X
plotData.X<-
  plotData.X %>% 
  rename(Vol="Vol.sum",
         BA="BA.sum",
         SPH="SPH.sum",
         QMD="QMD.mean")

# fit linear models
fit.vol<-lmer(Vol~BA.Target*Year+(1|Plot),data=plotData.X)
fit.ba<-lmer(BA~BA.Target*Year+(1|Plot),data=plotData.X)
fit.sph<-lmer(SPH~BA.Target*Year+(1|Plot),data=plotData.X)

# Summarize models
library(lmerTest)

# Create estimated marginal means
vol.means <- emmeans(fit.vol, ~ BA.Target*Year)
ba.means <- emmeans(fit.ba, ~ BA.Target*Year)
sph.means <- emmeans(fit.sph, ~ BA.Target*Year)

# Slope Contrasts
vol.contrast <- contrast(vol.means,interaction=c("pairwise","revpairwise"))
ba.contrast <- contrast(ba.means,interaction=c("pairwise","revpairwise")) 
sph.contrast <- contrast(sph.means,interaction=c("pairwise","revpairwise"))

# Mean contrasts
volM.contrast <- pairs(vol.means,by="Year")
baM.contrast <- pairs(ba.means,by="Year")
sphM.contrast <- pairs(sph.means,by="Year")

# CREATE TABLES

# create the data frame of ANOVA
standFit.X<-
  data.frame(Variable="Volume",anova(fit.vol, type="3")) %>% 
  rbind(data.frame(Variable="Basal area",anova(fit.ba, type="3"))) %>% 
   rbind(data.frame(Variable="Stand density",anova(fit.sph, type="3"))) %>%

  # formatting
  rownames_to_column("Fixed effect") %>% 
  dplyr::select(Variable,`Fixed effect`,Sum.Sq:F.value,P.value="Pr..F.") %>% 
  mutate_at(vars("F.value","P.value"),round,5) %>% 
  mutate(`Fixed effect`=rep(c("Treatment","Year","Trt*Year"),times=3)) %>% 
  rename(`Sum sq.`="Sum.Sq",
         `Mean sq.`="Mean.Sq",
         `Num. DF`="NumDF",
         `Den. DF`="DenDF",
         `F value`="F.value",
         `p value`="P.value")

# Create slope contrasts table
standFit.contrasts<-
  data.frame(Variable="Volume",vol.contrast) %>% 
  rbind(data.frame(Variable="Basal area",ba.contrast)) %>% 
  rbind(data.frame(Variable="Stand density",sph.contrast)) %>% 
  filter(Year_revpairwise %in% c("2019 - 1994")) %>% 
  dplyr::select(Variable,`Year contrast`="Year_revpairwise",`Treatment contrast`="BA.Target_pairwise",everything()) %>% 
  rename(`t ratio`="t.ratio",
         `p value`="p.value") %>% 
  mutate_at(vars(estimate:`t ratio`),round,3) %>% 
  mutate_at(vars(`p value`),round,5)

# Create mean contrasts table
standFitM.contrasts<-
  data.frame(Variable="Volume",volM.contrast) %>% 
  rbind(data.frame(Variable="Basal area",baM.contrast)) %>% 
  rbind(data.frame(Variable="Stand density",sphM.contrast)) %>% 
  dplyr::select(Contrast=contrast,Variable,Year,everything())

```

```{r plot comparisons at the species level, eval=FALSE}

# Summarize plot data at the species and treatment unit level
speciesData.X<-
  X %>%

  filter(Status=="Live") %>% 
  filter(Species %in% c("Spruce","Fir")) %>% 
    # Calculate volume
  left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  mutate(BA=d2ba(DBH)) %>% 
  dplyr::select(BA.Target,Plot:Species,Year,DBH,BA,vol,DBH.class,Height.Pred) %>% 
  mutate(Species=factor(Species)) %>%

  # Remove volumes below 18.5cm DBH
  mutate(vol=replace(vol,DBH<18.5,0)) %>%

  # now create plot sums grid so that you can calculate means
  group_by(BA.Target,Species,Plot,Year) %>% 
    summarise(Vol=sum(vol,na.rm=T)*pm,
              BA=sum(BA)*pm,
              SPH=n()*pm) %>%

  ungroup() %>%

  # Create QMD
  mutate(QMD.mean=sqrt(BA/(SPH*0.0000785)))

# fit linear model
# fit linear models
fit.vol.spp<-lmer(Vol~BA.Target*Year*Species+(1|Plot),data=speciesData.X)
fit.ba.spp<-lmer(BA~BA.Target*Year*Species+(1|Plot),data=speciesData.X)
fit.sph.spp<-lmer(SPH~BA.Target*Year*Species+(1|Plot),data=speciesData.X)

# Summarize models
library(lmerTest)

# Create estimated marginal means
vol.spp.means <- emmeans(fit.vol.spp, ~ BA.Target*Year*Species)
ba.spp.means <- emmeans(fit.ba.spp, ~ BA.Target*Year*Species)
sph.spp.means <- emmeans(fit.sph.spp, ~ BA.Target*Year*Species)

# Mean contrasts
volM.spp.contrast <- pairs(vol.spp.means,by="Year")
baM.spp.contrast <- pairs(ba.spp.means,by="Year")
sphM.spp.contrast <- pairs(sph.spp.means,by="Year")

```

### Rate of change from 1994 to 2019

All treatment units, including the control, increased volume, basal area and stand density over the 1994-2019 period (`r figNums("standPlots",display="cite")`).  Over this period, the low RBA treatment unit accrued basal area and density significantly faster (p<0.05) than the high RBA unit and control, and the high RBA unit accrued stand density faster than the control (p<0.05).  The rate of volume increment did not differ significantly between treatments and control.  Contrasts of estimated marginal means showed that basal area increment in the low RBA unit exceeded the high RBA and control unit by an estimated `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Basal area") %>%   pull(estimate) %>% round(.,1) %>% .[1]` 
and
`r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Basal area") %>%   pull(estimate) %>% round(.,1) %>% .[2]`
m² per hectare, respectively (p\<0.01 for both,
`r appNums("standFit.contrasts",display="cite")`).

In 1994, the low RBA unit had 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==1994) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
 `r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==1994) %>%   pull(estimate) %>%   abs() %>%   round(.,1)`
 less basal area (m² per hectare) than the high RBA unit and control, respectively, and by 2019, those differences had decreased to 
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:20") %>%  filter(Variable=="Basal area") %>%  filter(Year==2019) %>%  pull(estimate) %>%   abs() %>%   round(.,1)`
and
`r standFitM.contrasts %>%   filter(Contrast=="TU:10 - TU:Control") %>% filter(Variable=="Basal area") %>%   filter(Year==2019) %>%   pull(estimate) %>%   abs() %>%   round(.,1)` m² per hectare, respectively.  In terms of stand density, the low RBA unit accrued `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[1]` 
and
`r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[2]` more trees per hectare from 1994-2019, than the high RBA and control, respectively (p<0.05 for both), and the high RBA accrued  `r standFit.contrasts %>% as.data.frame() %>% filter(Variable=="Stand density") %>%   pull(estimate) %>% round(.,0) %>% .[3]` more trees than the control (p<0.05) over the same time period.  By 2019, there were no significant stand density differences between the treatments and control.

The rate of change varied by measurement period.  In the low RBA unit, basal area and volume increment peaked at 
`r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="1997-2009") %>% pull("Vol") %>%   round(.,1)` m³/ha, respectively, in the 1997-2009 period; increments declined to `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:10") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively in the subsequent measurement period of 2009-2019.  In the high RBA unit, periodic annual basal area and volume increment increased with each measurement period and peaked in the 2009-2019 measurement period at `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:20") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively.  Conversely, basal area and volume increment in the control unit was highest in the 1994-1997 period, and declined with each subsequent measurement period to `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("BA") %>%   round(.,2)` m²/ha and
 `r perInc.X %>% filter(BA.Target=="TU:Control") %>% filter(Period=="2009-2019") %>% pull("Vol") %>%   round(.,1)` m³/ha respectively..

### Growth dominance coefficients

```{r calculate growth dominance}

# Filter only for trees that have measurements in 1994 and 2019 
gd.treeList<- # assign variable
  X %>% 
  filter(TreeID!="New") %>% # remove new trees
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(Status=="Live") %>% # filter for live trees
  group_by(TreeID) %>% 
  summarise(N=n()) %>% 
  filter(N>=2) %>% #filter for trees with at least 2 measurements
  pull(TreeID)

# derive periodic increment (period diameter or basal area increment)
gd.BAInc<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  mutate(BA=d2ba(DBH)) %>% # convert DBH to basal area
  dplyr::select(Plot,TreeID,Species,Year,BA) %>% 
  pivot_wider(names_from=Year,values_from=BA) %>% 
  mutate(PBAI_2019=`2019`-`2009`,
         PBAI_2009=`2009`-`1997`,
         PBAI_1997=`1997`-`1994`,
         PBAI_1994=`1994`-`1992`) %>% 
  dplyr::select(Plot,TreeID,Species,starts_with("PBAI_"))

gd.Inc<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  dplyr::select(Plot,TreeID,Species,Year,Vol=vol) %>% 
  pivot_wider(names_from=Year,values_from=Vol) %>%
   mutate(PVI_2019=`2019`-`2009`,
         PVI_2009=`2009`-`1997`,
         PVI_1997=`1997`-`1994`,
         PVI_1994=`1994`-`1992`) %>% 
  dplyr::select(Plot,TreeID,Species,starts_with("PVI_")) %>% 
  inner_join(gd.BAInc,by=c("Plot","TreeID","Species")) %>% 
  pivot_longer(
    cols = -c("Plot","Species","TreeID"),
    names_to = c("Var", "Year"),
    names_sep = "_",
    values_to = "Inc")

gd.prelim<-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  dplyr::select(Plot,TreeID,Species,Year,DBH.Init=DBH) %>% 
  mutate(Year=fct_recode(Year,tmp_1994="1992",
                         tmp_1997="1994",
                         tmp_2009="1997",
                         tmp_2019="2009")) %>% 
  mutate(Year=str_remove(as.character(Year),pattern="tmp_")) %>% 
  mutate(Year=factor(Year)) %>% 
  mutate(BA.Init=d2ba(DBH.Init)) %>% 
  dplyr::select(Plot:Year,BA.Init)

gd.Dat <-
  X %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% gd.treeList) %>% 
  dplyr::select(Plot,TreeID,Species,Year,Status,DBH,Height) %>% 
  left_join(gd.Inc,by=c("Plot","TreeID","Species","Year")) %>% 
  left_join(gd.prelim,by=c("Plot","TreeID","Species","Year")) %>% 
  filter(Status=="Live") %>% 
  drop_na(Inc)

# clean up some objects
  rm(gd.prelim,gd.BAInc,gd.Inc)

# # sgr.Dat
#   sgr.Dat<-
#     gd.Dat %>%
#     filter(Var=="PBAI") %>% 
#     ungroup() %>% 
#     group_by(Plot) %>% 
#     mutate(Stand.Inc=sum(Inc),
#            Stand.BA=sum(BA.Init)) %>% 
#     mutate(Prop.growth=Inc/Stand.Inc,
#            Prop.size=BA.Init/Stand.BA) %>% 
#     left_join(plotDat,by="Plot")

# Plots of GD

  ## Basal area
  gd.cumu.Inc<-  
    gd.Dat %>% 
    filter(Var=="PBAI") %>% 
    arrange(Plot,Year,BA.Init) %>% 
    group_by(Plot,Year) %>%

    # CUmulative increment
    mutate(Cumu.Inc=cumsum(Inc)) %>% 
    mutate(Cumu.Inc2=max(Cumu.Inc)) %>% 
    mutate(Cumu.Prop.Inc=Cumu.Inc/Cumu.Inc2) %>% 
    dplyr::select(-Cumu.Inc2) %>%

    # Cumulative stand basal area
    mutate(Cumu.BA=cumsum(BA.Init)) %>% 
    mutate(Cumu.BA2=max(Cumu.BA)) %>% 
    mutate(Cumu.Prop.BA=Cumu.BA/Cumu.BA2) %>% 
    dplyr::select(-Cumu.BA2) %>%

    # Bring in plot data
     left_join(plotDat,by="Plot")

    # Calculate Growth Dominance statistic
    gd.stats<-
      gd.cumu.Inc %>% 
      mutate(GDS=(Cumu.Prop.Inc+lag(Cumu.Prop.Inc,default=0))*(Cumu.Prop.BA-lag(Cumu.Prop.BA,default=0))) %>% 
      ungroup() %>% 
      group_by(Plot,Year) %>% 
      summarise(GDS=1-sum(GDS)) %>% 
      left_join(plotDat,by="Plot") %>% 
      ungroup() %>% 
      group_by(BA.Target,Year) %>% 
      summarise(GDS.mean=round(mean(GDS),3),
                GDS.sd=round(sd(GDS),2)) %>% 
      mutate(GDS.label=paste0("GDS: ",GDS.mean," (",GDS.sd,") "))

  #   gd.stats.plot<-
  #     gd.stats %>% 
  #   filter(Year %in% c(2009,2019)) %>% 
  # ggplot()+
  # aes(x=BA.Target,y=GDS,fill=Year)+
  # geom_boxplot()
  #

  # Plotting
  # ## Growth Dominance statistic
  gd.plot<-
    gd.cumu.Inc %>%
    filter(Year %in% c(2009,2019)) %>%
    ggplot()+
    aes(x=Cumu.Prop.BA,y=Cumu.Prop.Inc,group=BA.Target)+
    geom_point(colour="grey")+
    geom_smooth(se=F,size=1.5)+
    geom_abline(slope=1,intercept=0,linetype = "dashed",size=1.5)+
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    facet_grid(Year~BA.Target,scales="free_y")+
    xlab("Cumulative proportion of stand size (basal area)")+
    ylab("Cumulative proportion of stand growth (basal area)")+
    facet_grid(Year~BA.Target)+
    geom_text(data=filter(gd.stats %>% filter(Year %in% c(2009,2019))),
            size=4,
            mapping = aes(x = Inf, y = 0.1, label = GDS.label),
            hjust="right")

  ## Plot measurement year GSD
      # gd.stats %>% 
      #   ggplot()+
      #   aes(x=Year,y=GDS,group=Plot,color=Plot)+
      #   geom_line()+
      #   facet_wrap(~BA.Target)
      #

  ## GSize-Growth relationship statistic
      # sgr.ind.plot<-
      #   sgr.Dat %>% 
      #  ggplot()+
      #   aes(x=Prop.size,y=Prop.growth,group=Plot,color=BA.Target)+
      #   geom_point()+
      #   geom_smooth(aes(group=Plot),method="lm",se=F,color="black")+
      #   facet_wrap(~Year)+
      #       theme(aspect.ratio = 1)+
      #     facet_grid(BA.Target~Year)

```

The growth dominance coefficient was negative for all plots and time periods, indicating that smaller trees contributed
proportionally more to stand basal area growth than larger trees
(`r figNums("gds.plot",display="cite")`). Over the last two measurement
periods of 1997-2009 and 2009-2019, the growth dominance coefficient increased in all plots except
one (not shown), indicating that individual tree growth contributions to
stand growth became more proportional to tree size over the last 11
growing seasons.

### Stand attributes by diameter class

Among treatments and throughout the stand development after treatment,
the stands maintained a negative exponential (also referred to as
'reverse-J') diameter class distribution
(`r figNums("diamK.plot",display="cite")`).
Immediately after the partial harvest, the diameter class distribution
of volume differed among treatments. In the 27 years since partial
harvest, volume in all three treatment units was mostly concentrated in
the 22.5-27.5cm DBH class, and mostly comprised of fir
(`r figNums("diamVol.plot",display="cite")`). Over both time periods
shown in the figure and all three treatments, most of the spruce volume
was concentrated in DBH classes spanning 37.5 to 47.5cm. Basal area
trends were very similar to volume (not shown).

Q factor
```{r compute Q factor}

# Fit negative exponential distribution to plots/time combinations, and then combine
source(here::here("manuscript","negExp.R"))

# compute for each plot and time period
x.qFact<-
  X %>% 
  mutate(DBH.class=as.numeric(as.character(DBH.class))) %>% 
  group_by(BA.Target,Plot,Year,DBH.class) %>% 
  summarise(k=n()) %>% 
  complete(BA.Target,Plot,Year,DBH.class,fill=list(k=0)) %>% 
  group_by(BA.Target,Year,DBH.class) %>% 
  summarise(k=mean(k)) %>% 
  mutate(k=k*20) %>% 
  filter(k>0)

# function to apply negExp to each year
nexExp.year<-function(year){

  x<-
    x.qFact %>% 
    filter(BA.Target=="TU:10") %>% 
    filter(Year==year)

    negExp(x$DBH.class,y=x$DBH.class)

  }

}

```

### Species composition

```{r generate species composition data}

sppComp.X<-
  X %>% 
  filter(Status=="Live") %>% # filter for live trees
  filter(Species %in% c("Spruce","Fir")) %>% # filter for spruce and fir trees

  # Compute basal area and volume
 left_join(filter(volCoef,VolType=="total"),by="Species") %>% 
  mutate(vol=exp(1)^b0*DBH^b1*Height.Pred^b2) %>% 
  mutate(BA=d2ba(DBH)) %>% 
  dplyr::select(BA.Target,Plot:Species,Year,DBH,BA,vol) %>% 
  mutate(Species=factor(Species)) %>%

  # Need to add zero values for spruce in Plot 19
  add_row(BA.Target="TU:10",Plot=19,Year=c(1992,1994,1997,2009,2019),Species="Spruce",DBH=20,vol=0,BA=0) %>%

  # Remove volumes below 18.5cm DBH
  mutate(vol=replace(vol,DBH<18.5,0)) %>%

  # summarise 
  group_by(BA.Target,Plot,Species,Year) %>% 
  summarise(sph=n()*pm,
            BA=sum(BA)*pm,
            Vol=sum(vol)*pm) %>% 
  ungroup() %>%

  # pivot_longer
  pivot_longer(cols=sph:Vol,names_to="Var",values_to = "Value")

plotTotal.X<-
  sppComp.X %>% 
  pivot_wider(names_from="Var",values_from = "Value") %>% 
  group_by(BA.Target,Plot,Year) %>% 
   summarise(sph=sum(sph),
            BA=sum(BA),
            Vol=sum(Vol)) %>% 
  pivot_longer(cols=sph:Vol,names_to="Var",values_to="Plot.Total") %>% 
  inner_join(sppComp.X,by=c("BA.Target","Plot","Year","Var")) %>%

  # create proportion column
  mutate(Prop=Value/Plot.Total) %>% 
  ungroup() %>%

  # a Few plots have no values in some years
  drop_na(Prop)

plotTotal.table<-
  plotTotal.X %>% 
  group_by(BA.Target,Year,Species,Var) %>% 
  summarise(Prop=mean(Prop,na.rm=T))

```

A comparison of proportional species composition by diameter class,
treatment and year showed that fir and spruce volume proportion by
diameter class varied similarly between the high RBA and control units
(`r figNums("sppCompDiam.plot",display="cite")`). In 1994, the largest
diameter classes were dominated by spruce, whereas fir was the dominant
species in the smallest diameter classes. This pattern was also found
for basal area and tree density (not shown). By 2019, the fir-spruce
composition by distribution class was maintained in these two units,
with small increases in spruce composition in the smallest diameter
class.

In the low RBA, the species composition distribution by diameter class
was different. Here, both fir and spruce maintained similar proportions
across diameter classes smaller than 37.5cm DBH. Spruce formed the
leading species in diameter classes above 37.5cm DBH, with the exception
of the largest class.

## Mortality

```{r mortality data}

# Create dead dataset

# Dead species totals
deadList<-
  X %>% 
  filter(Status=="Dead") %>% 
  mutate(ID=paste(Plot,TreeID,sep="-")) %>% 
  group_by(ID) %>% 
  summarise_all(first)

# Summarize mortality by treatment unit
dead<-
  X %>% 
  filter(Status=="Dead") %>% 
  mutate(ID=paste(Plot,TreeID,sep="-")) %>% 
  mutate(Species=fct_recode(Species,`Douglas-fir`="Fd")) %>% 
  group_by(ID) %>% 
  summarise_all(first) %>% 
  ungroup() %>% 
  mutate(BA=d2ba(DBH)) %>% # convert to basal area per hectare 
  group_by(BA.Target,DBH.class,Year) %>% 
  summarise(BA=sum(BA),
            k=n())

# Summarize mortality by plot - this will be used for Figure
deadPlot.X<-
  X %>% 
  filter(Status=="Dead") %>% 
  # filter(Species=="Spruce") %>% #dot this if you want to summarize by species
  mutate(ID=paste(Plot,TreeID,sep="-")) %>% 
  mutate(Species=fct_recode(Species,`Douglas-fir`="Fd")) %>% 
  group_by(ID) %>% 
  summarise_all(first) %>% 
  ungroup() %>% 
  mutate(BA=d2ba(DBH)) %>% # convert to basal area per hectare 
  group_by(BA.Target,Plot) %>% 
  summarise(BA=sum(BA),
            k=n()) %>% 
  full_join(plotData.X,by=c("BA.Target","Plot")) %>% 
  filter(Year==1994) %>% 
  rename("BA.dead"=BA.x,"Num.dead"=k,"BA"=BA.y) %>% 
  mutate(BA.dead=replace_na(BA.dead,0),
         Num.dead=replace_na(Num.dead,0)) %>% 
  mutate(prop.K=Num.dead/SPH*100,
         prop.BA=BA.dead/BA*100)

# linear model to analyze mortality by plot volume
deadPlot.ANOVA<-
  lm(Num.dead~BA,data=deadPlot.X)

dead.Summ<-
  dead %>% 
  ungroup() %>% 
  group_by(BA.Target) %>% 
  summarise(BA=sum(BA)/sum(dead$BA)*100,
            k=sum(k)/sum(dead$k)*100) %>% 
  mutate_if(is.numeric,round,1)

```

A total of `r deadList %>% nrow()` trees
(`r deadList %>% filter(Species=="Fir") %>% nrow()` fir and
`r deadList %>% filter(Species=="Spruce") %>% nrow()` spruce) died in
`r length(unique(deadList$Plot))` plots from 1994 to 2019. The number of
trees that died per plot from 1994 to 2019 increased significantly with
plot basal area in 1994 (`r figNums("mort.plot",display="cite")`, p \<
0.0001). In the majority of cases, we could not identify a causal agent, and overall, mortality was generally not attributable to a particular forest
health factor. Linear models fitted separately for fir and spruce
showed that fir mortality had a similar relationship to plot basal area,
whereas spruce mortality did not (not shown). The diameter distribution
of fir mortality followed a negative exponential distribution, with
mortality concentrated in smaller diameter classes
(`r figNums("mortSppDiam.plot",display="cite")`). Spruce mortality was
concentrated in larger diameter classes. By treatment unit, the control
unit had the highest mortality, whereas the high-removal treatment unit
had the lowest mortality (`r figNums("mortDiam.plot",display="cite")`).
The majority of mortality occured during the most recent measurement
period of 2009-2019.

# Discussion

## Uneven-aged stand management
The successful application of partial harvesting in uneven-aged, multi-species stands can be complicated [@smith1997practice], especially where the objectives include the sustained production of wood over multiple generations of trees.  In this context, partial harvesting seeks to re-allocate growing space and resources to trees across a range of species and size classes to ensure their survival and productivity.  Knowledge gleaned from thinning treatments in even-aged stands can help inform silviculture practices in more complex stands, however, complexities associated with competition between size classes and species requires additional consideration.

## Tree- and stand-level growth response to partial harvest

### Species

At the tree-level, our results suggest that while both hybrid white
spruce and subalpine fir respond to thinning with faster radial growth,
heavy thinning can elicit a stronger radial growth response in spruce,
compared to fir. Spruce trees in the low-removal harvest treatment did
not respond with a significant growth release. Thus, our results suggest
that a relatively high removal of the stand will result in increased
spruce growth, and this release will exceed that of co-occurring fir
trees. Other studies have shown that residual spruce growth responds
favourably to heavy thinning. A 10-year analysis of white spruce growth
underneath an aspen canopy found that a complete or high overstory
removal resulted in immediate and significant spruce radial and height
growth responses, compared to control and partial cutting treatments
[@smith2016sprucing]. In a 35-year old white spruce stand, diameter
increment was maximized with the heaviest thinning intensity, however,
the treatment thinning from below [@Stiell1980]. Frank [-@frank1973course] also
showed that basal area increment in a 70-year old white spruce stand increased with the
amount of thinning. Competition for light is a key factor that limits
white spruce growth, and opening up the stand to allow for increased
light penetration may be a primary mechanism that enhances white spruce
growth responses [@Lieffers1999]. High basal area removal may also
increase air and soil temperatures, which can preferentially favour
spruce growth responses compared to fir [@Lajzerowicz2004]. The
relatively low spruce growth response in the low-removal treatment may
be explained in that the gaps created may have been too small to elicit
as strong a growth response as observed in the high-removal thinning
treatment. Light levels may differ little between forest canopies that
are closed or contain small gaps, and strong vertical
light gradients may result in relatively little light regime changes for
understorey trees [@Bartemucci2002].

In comparison, both partial harvest treatments stimulated a growth
response in fir, compared to the control. Firs are considered to be better adapted than spruce to
low light conditions in the understory, allowing them to persist longer
in undisturbed forest [@Antos2000] and potentially able to respond
faster to even modest changes in understory light [@Dumais2014; @Messier1999]. Although fir growth may respond at lower levels of
overstorey removal, the growth response potential of fir may be lower
than spruce. Comparative studies of spruce and fir growth rates present
differing evidence regarding genus-level differences in release potential to
thinning. In mature Engelmann spruce (*Picea engelmannii*) - subalpine
fir forests in western North America, understory spruce growth rates
exceeded those of fir, allowing spruce to achieve canopy position
faster, including after canopy gaps were created through disturbance
[@Antos2000; @Andrus2018]. In a mature forest in central BC,
spruce growth exceeded that of subalpine fir after age 80 [@eis1983regeneration]. Because spruce tends to live longer than subalpine
fir, higher growth rates at later ages may allow this species to become
larger than fir. In contrast, thinning studies in spruce-fir forests in
eastern North America report that balsam fir (*Abies balsamifera*)
responds to release with faster height and diameter growth than
co-occurring red spruce (*Picea rubens*) [@Dumais2008; @Dumais2008a; @Dumais2014; @Messier1999]. Other studies have found no
difference in growth response between spruce and fir species after
treatment [@mccaughey1982understory]. Our study also suggests that spruce and fir growth responses to lower levels of basal area removal are
similar.

Post-treatment growth rates increased with tree size at the time of
treatment, for both species. This has been found in other studies as
well (references) and may be related to size-asymmetric competition
whereby larger trees preempt resources, especially light, from smaller
individuals [@Bose]. Thinning treatments may exacerbate
size-assymetric competition [@cordonnier2015gini], and growth increases from
treatment will be predominantly concentrated in the larger tree size
classes [@Bose; @schwinning1998mechanisms]. Smaller trees will likely continue
to experience competition and will not benefit as much from the thinning
treatment. However, thinning treatments may not promote growth responses in the
largest dominant trees in a stand either, as these trees may not
experience significant changes in resource availability or may have
already achieved maximal growth rates [@ryan1997age, @Bose].

Although growth responses were concentrated in the larger trees, our analysis of growth dominance indicated that smaller trees contributed proportionally more to stand growth than larger trees.  Binkley et al. [-@Binkley2006] propose that GDC in developing forests increases from 0 to positive during early development, reflecting the growth dominance of larger robust trees as canopy closure is achieved.  As the forests age, the GDC may switch to negative, reflecting several co-occuring processes.  Older trees may start to senesce, with growth reductions and increased mortality.  Concurrently, younger trees may increase their resource use efficiency, and/or increase their allocation to leaf biomass [@baret2017long], partially in response to increasing resource availability resulting from the senescence of older trees.  Data from old-growth Engelmann spruce – subalpine fir forests in Colorado [@Binkley2006] as well as old-growth boreal mixedwood stands [@baret2017long] support this proposed model of growth dominance development.  While we found a similar negative growth dominance in the stands of this study, a comparison between the two recent measurement periods showed that growth dominance actually increased over time (i.e., tree growth became increasingly proportional to tree size).  This finding is unique from other studies, and suggests that either larger trees are increasing their growth rates relative to smaller trees, or smaller trees are beginning to experience increased competition and may be experiencing lower growth rates.  Further data from these stands will help evaluate the long-term trend in GDC.

Our results suggest that thinning did not substantially alter the growth dominance of the stand at any point 27 years after treatment.  Similar results were reported after thinning in uneven- and even-aged hardwood stands in eastern Canada [@Lemire2020].  While thinning treatments stimulated a growth response in all trees, the increase in growth was more or less proportional to tree size.  A significant decrease in GDS could be potentially achieved in older stands where the treatment seeks to release suppressed larger trees [@Lemire2020].  Conversely, thinning even-aged red pine (*Pinus resinosa*)  stands reduced growth dominance compared to unthinned stands, regardless of thinning type, presumably by allocating more growing space to previously suppressed trees [@Bradford2010].  The use of GDC to evaluate the effects of partial cutting may be insufficient on its own, and should be used in conjunction with tree-level analyses [@Lemire2020].  Forest managers should account for potential size differential when planning thinning
treatments. For example, a silviculture treatment that seeks to release
a suppressed understorey may have to ensure that sufficient overstorey
trees are removed to reduce assymetric competition and growth dominance after treatment.

## Stand-level responses to thinning treatments

The thinned treatments increased tree density faster than
the control, and within three decades, there were no significant density
differences between the two harvested treatments and control. This
likely resulted from two processes; growth and mortality. Increasing
stand density reflects the growth of smaller trees that achieved the
minimum diameter class (i.e., 7.5cm DBH) in this study. The faster
increases in stand density of the thinned treatments compared to the
control likely reflect that smaller trees in the thinned stands
responded with faster diameter increment after treatment (as previously
discussed). We note that our study did not account for new seedlings
that may have become established over the period of the study. Based on
increasing densities in the smaller diameter classes, these stands will
likely maintain a negative-exponential diameter distribution, which is
considered typical of older, uneven-aged stands in this region.

Stand densities between all units likely also converged because of
different mortality rates. Tree mortality increased with stand-level
basal area, and the thinned stands had the lowest mortality over the
time of the study. Most of the mortality was concentrated in fir trees
among the smaller size classes. Density-dependent tree mortality is
common in younger, even-aged stands, reflecting intertree competition
for light and other resources, and resulting in increasing uniformity of
tree distribution [@Larson2015]. In the uneven-aged stands of our study,
density-dependent mortality likely reflects indicate the effects of
assymetric competition and suppression of . In the unthinned stands,
smaller trees died more frequently as a result of suppression from the
larger trees, whereas thinning reduced this competition.
Density-dependent mortality in uneven-aged, mixed-species forests can
also vary with drought and site-level conditions such as aspect and soil
moisture regime, which needs to be considered when planning stand
thinning treatments [@VanGunst2016].

Species composition

The low-removal thinning treatment also maintained a species composition
by diameter class similar to the unthinned stands. In these units,
subalpine fir was the leading species by frequency in the smaller
diameter classes, and spruce was the leading species in the larger
classes. This is a stand structure and composition common at low- to
mid-elevations in the interior of British Columbia. In closed forests
with undisturbed forest floors, subalpine fir can have higher seedling
establishment success than spruce, leading to higher densities of fir in
the understorey. Compared to spruce, subalpine fir can be relatively
short-lived, and can die after 250 years in the canopy. Spruce can live
for over 400 years and thus can achieve larger sizes and dominant canopy
positions.

The high-removal unit had a somewhat different species composition by
diameter class. Here, fir was the leading species in all diameter
classes. This means that the high-removal thinning treatment converted
the stand composition to fir-leading in all tree sizes. This result may
be concern for forest managers, as spruce is often preferred as a crop
species in British Columbia and elsewhere [@Deal]. However, the stands in this treatment had
more spruce in the smaller diameter classes than the other treatment
units and over the 27-year period of this study, spruce monotonically
increased its composition (by volume and basal area) compared to fir.
Over time, we expect that spruce will continue to increase its
proportion relative to fir and eventually achieve a similar species
composition - diameter distribution as the other units.

### Tree mortality
Our results suggest that mortality increased with stand density.  The majority of trees that died were fir, and these were predominantly concentrated in the smaller size classes.  These findings concur with other studies that found that risk or tree mortality increases with stand density, competition and tree-level factors such as live crown and growth [@Valkonen2020; @VanGunst2016].  Competition from overstory trees is likely a contributing cause to most of the mortality recorded in our study.    Because we did not record trees below 7.5cm DBH, we cannot quantify total tree mortality in the stand.  Given the relatively low mortality in the thinning treatments, it does appear that partial harvesting may effectively capture some volume that would otherwise be lost to mortality, potentially affecting the total net volume recovered from a stand.  Density-dependent mortality may vary with other factors such as drought [@VanGunst2016], and climate change may alter competitive interactions between species and size classes.  We also note that as of 2019, there was a large spruce bark beetle (*Dendroctonus rufipennis*) outbreak in the study region; presumably, the larger spruce trees in this study may be removed by this forest health agent in the short-term.  This may provide an opportunity for further study into stand responses after mortality of larger stems.

### Stand-level growth responses

Over the 27-year period of this study, plots with high-removal harvest
treatments accrued basal area significantly faster than the control and
low-removal harvest treatment, likely due to increases in tree-level
growth and stand density discussed previously. As a result, differences
in basal area between the high-removal harvest unit and the control
reduced from ca. 17m2/hectare in 1994 to ca.9m2/hectare in 2019.

In contrast, basal area increment between the low-removal unit and the
control did not differ significantly over the 1994-2019 period, and the
difference between these two units remained relatively constant over all
measurement periods at ca. 10m2/hectare. The similar rates of basal area
increment between the low-removal thinning treatment and the control are
difficult to explain. Tree-level results showed that fir grew faster in
the low-removal unit than the control, but spruce did not. As well, by
2019 there were no significant differences in stand density between the
two units. Thus, it appears that while the fir trees may have increased
growth as a response of low-removal, the growth increases were
insufficient to cause a significant change in basal area increment.
Further monitoring in this study will help determine the future
trajectory of the low-removal thinning treatment.

While our study found that high-removal partial harvest resulted in
faster basal area growth, we did not find evidence that thinning changed
volume increment in trees greater than 17.5cm DBH. A possible
explanation is that tree growth responses to thinning might be primarily
concentrated in radial growth, as opposed to height growth; this would
be important because the volume formula used in this study is based on
both tree diameter and height [@nigh2016]. We note that it is often
difficult to precisely capture height growth changes in mature trees
using conventional field-based methods [@vepakomma2011response], which makes
capturing height changes problematic in these sorts of studies. We also
utilized a 17.5cm DBH limit for volume calculations; as such, our volume
measures only reflect the contribution of the larger trees in the stand,
and do not include volume increment in smaller trees.

The partial harvest treatment appeared to change the timing of peak
stand growth. In the control unit, volume and basal area increment
peaked in the 1994-1997 measurement period, and declined thereafter,
with the strongest decline in annual growth over the last measurement
period of 2009-2019. Age-related declines in stand productivity may be
associated with decreasing water-use efficiency and reduced
photosynthesis, with the age of maximal growth rate varying with factors
such as species and site conditions [@ryan1997age]. In both thinned
treatments, maximal growth rates peaked in the 1997-2009 period. Thus,
it appears the thinning treatment may have delayed the period of maximal
growth rates by approximately 10 years. The timing of thinning
treatments can be important. Stands that are past their period of
age-related maximal growth rate may not have as much capacity to respond
to treatments as a younger stand, although this is likely a more
important factor for shade-intolerant species rather than the spruce and
fir trees in this study [@Bose]. At the time of this study, the
high-removal thinning treatment had maintained an annual basal area and
volume increment that was similar to previous measurement periods; if
this high level of increment continues, differences in basal area and
volume will continue to diminish compared to the control, and this stand
may exceed the low-removal thinning treatment.

## Management of uneven-aged stands

## Management considerations

1.  Both species responded with increased growth after partial harvest.
    High removal of basal area in these forests stimulated stronger
    spruce growth. This information can help silviculturists manipulate
    stand structure to preferentially ...
2.  Partial harvesting at both intensities delayed peak volume and basal
    area increment by approximately 10 years. Annual increment in the
    control unit declined faster over the last measurement period
    compared to previous periods, perhaps suggesting that the stand is
    becoming mature and losses due to mortality may substantially offset
    growth increment.
3.  GIven that larger trees responded with faster growth than smaller
    trees, silviculturists need to consider the implications of this on
    residual stand structure and response after treatment.

\newpage

# Tables

`r tableNums("fitBA.table",caption="Summary of two mixed effects models of tree-level basal area increment")`

```{r summarize tree level BAI model}

fitBA.random<-
  fitBA %>% 
  VarCorr() %>% 
  as_tibble() %>% 
  bind_rows(as_tibble(VarCorr(fitBApct))) %>% 
  dplyr::select(Group=grp,Var.=vcov,St.Dev=sdcor)

fitBA.fixed<-
  fitBA %>% 
  anova() %>% 
  tidy() %>% 
  bind_rows(tidy(anova(fitBApct))) %>% 
  mutate(term=c("TU*Tree Size*Sp.")) %>% 
  mutate(p.value=c("<0.00001","<0.00001")) %>% 
  mutate(R2=c(round(r.squaredGLMM(fitBA)[2]*100,1),
              round(r.squaredGLMM(fitBApct)[2]*100,1))) %>% 
  add_column(Model=c("BAI","BAI (%)"),.before="term") %>% 
  add_row(.after=1) %>% 
  add_row(.after=3) %>% 
  cbind(fitBA.random) %>%

  # format table
  setNames(c("Model","Term","Sum.Sq.","Mean.Sq.","Num.DF","Den.DF",
             "Stat.","P value","R²","Group","Var.","St.Dev.")) %>% 
  mutate_at(vars(Sum.Sq.:Stat.),round,3) %>% 
  mutate_at(vars(Var.:St.Dev.),round,3)

# set typology
typology <- data.frame(
  col_keys=names(fitBA.fixed),
  what = c( "Model",rep("Fixed effects",8),rep("Random effects",3)),
   what=names(fitBA.fixed),
  stringsAsFactors = FALSE )

fitBA.fixed %>%

  flextable() %>%

  set_header_df(mapping = typology, key = "col_keys" ) %>% 
  colformat_num(digits = 2, na_str = "") %>% 
  colformat_num(j=9,digits = 0, na_str = "") %>% 
  merge_h(part = "header") %>% 
  # merge_v(part = "header") %>%

  valign(valign="top",part="all") %>% 
   hline(i=c(2,4),border=fp_border(color="black")) %>% 
     hline(part="header",border=fp_border(color="black")) %>% 
  vline(j=c(1,9),part="all",border=fp_border(color="black")) %>%  
 align(align="left",part="body") %>%

  fontsize(size=7,part="all") %>% 
    fix_border_issues() %>% 
  autofit()

```

*The dependent variable in the first model (BAI) is tree basal area
increment from 1994-2019. The dependent variable in the second model
(BAI (%)) is basal area increment from 1994-2019 as a percentage of the
tree's basal area in 1994.*

\newpage

`r tableNums("standFit.X",caption="Output from three separate linear models testing the effect of the interaction between treatment and year on plot-level volume, basal area and density." )`

```{r stand variable model output}

standFit.X %>% 
  flextable() %>% 
  align(align="left",part="all") %>% 
  merge_v(j="Variable") %>% 
  valign(valign="top",part="body") %>% 
    hline(i=c(3,6,9),border=fp_border(color="black")) %>% 
   fontsize(size=7,part="all") %>% 
    fix_border_issues() %>% 
  autofit()
```

*Trt, treatment; Sq., squares, Num., numerator; Den., denominator; DF,
degrees of freedom*

\newpage

# Figures

```{r tree BAI by treatment species and diameter class}

# Create plot of periodic increment

# Using estimated marginal means
  emmeans(rg,~Init|BA.Target+Species,type="response") %>% 
  as.data.frame() %>%
  mutate(Init=ba2d(Init)) %>% # convert initial basal area back to diameter

  # plotting
  ggplot()+
  aes(x=Init,y=response,fill=Species,color=Species)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(0.05))+
  xlab("1994 DBH (cm)")+
  ylab("Tree BAI from 1994-2019 (m²)")+
  facet_wrap(~BA.Target)+
  theme(legend.position="bottom")

```

`r figNums("treeGrowth.plot",caption="Estimated marginal means of spruce and fir basal area increment (m²) across five inital DBH classes and three treatments.  Whiskers are standard error of the mean.  Responses are back-transformed from the model.  BAI, basal area increment")`

\newpage

```{r tree BAI as percentage by treatment species and diameter class}

# Create plot of periodic increment

# Using estimated marginal means
  emmeans(rg.pbaipct,~Init|BA.Target+Species,type="response") %>% 
  as.data.frame() %>%
  mutate(Init=ba2d(Init)) %>% # convert initial basal area back to diameter

  # plotting
  ggplot()+
  aes(x=Init,y=response,fill=Species,color=Species)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(0.05))+
  xlab("1994 DBH (cm)")+
  ylab("Tree BAI from 1994-2019 (% increase)")+
  facet_wrap(~BA.Target)+
  theme(legend.position="bottom")

```

`r figNums("treeGrowthPct.plot",caption="Estimated marginal means of spruce and fir basal area increment (expressed as a percentage increase relative to 1994 DBH) across five inital DBH classes and three treatments.  Whiskers are standard error of the mean.  Responses are back-transformed from the model. BAI, basal area increment")`

\newpage

```{r plot means plots, fig.height=10,eval=FALSE}

# Plotting

standData.X %>% 
  setNames(c("BA.Target","Year","Vol_mean","Vol_se",
             "BA_mean","BA_se",
             "SPH_mean","SPH_se",
             "QMD_mean","QMD_se")) %>%

  # Need to change year to numeric for plotting
  mutate(Year=as.numeric(as.character(Year))) %>%

  # reshape for ggplot
  pivot_longer(cols=Vol_mean:last_col(),names_to=c("Var","Metric"),names_sep="_",values_to="Values") %>%   pivot_wider(names_from="Metric",values_from="Values") %>%

  # Filter to remove quadratic mean diameter
  filter(Var!="QMD") %>%

  # Reset factor levels for plotting
  mutate(Var=fct_relevel(Var,"Vol","BA","SPH")) %>% 
  mutate(Var=fct_recode(Var,Volume="Vol",
                        `Basal area`="BA",
                        `Stand density (stems per hectare)`="SPH")) %>%

  ggplot()+
  aes(x=Year,y=mean,fill=BA.Target,color=BA.Target,group=BA.Target)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, width=.2))+
                   # position=position_dodge(.9))+
   scale_x_continuous(limits=c(1994,2019),breaks=c(1994,1997,2009,2019),
                     minor_breaks = c(1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis
  facet_wrap(~Var,scales="free_y")+
  xlab("Plot mean")+
  theme(legend.position = "bottom")

# Old code to produce separate grapohs below

# Create function
# standPlot<-function(standVar,y.label) {
# 
#     standData.X %>% 
#     mutate(Year=as.numeric(as.character(Year))) %>% 
#     dplyr::select(BA.Target,Year,standVar,paste0(standVar,".se")) %>% 
#     setNames(c("BA.Target","Year","Mean","SE")) %>% 
#   
#     ggplot(aes(x=Year,y=Mean,color=BA.Target,group=BA.Target))+
#     geom_point()+
#     geom_line()+
#     geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE, width=.2),
#                    position=position_dodge(.9))+
#     # facet_wrap(~BA.Target)+
#       scale_x_continuous(limits=c(1992,2019),breaks=c(1992,1994,1997,2009,2019),
#                      minor_breaks = c(1992,1994,1997,2009,2019),
#                      name="Measurement year")+ # stretch out x-axis
#   theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
#         legend.position = "none")+
#     ylab(y.label)
# 
# }
#     
# standVol.plot<-
#   standPlot("Vol",y.label="Vol. (m³/ha)")+
#   theme(axis.text.x=element_blank(),axis.title.x=element_blank())
#   
# standBA.plot<-
#   standPlot("BA",y.label="BA (m²/ha)")+
#   theme(axis.text.x=element_blank(),axis.title.x=element_blank())
#   
# standK.plot<-
#   standPlot("SPH",y.label="Tree dens. (sph)")
#   
# standQMD.plot<-standPlot("QMD",y.label="QMD (cm)") 
# 
# standPlots<-cowplot::plot_grid(standVol.plot,standBA.plot,standK.plot,ncol=1,
#                       align="v")
# 
# # display plot
# standPlots
```

```{r plot means by species, fig.height=7}

plotSums<-
  plotTotal.X %>% 
  group_by(BA.Target,Plot,Year,Var) %>% 
  summarise(Value=sum(Value)) %>% 
  mutate(Species="Total") %>% 
  dplyr::select(BA.Target,Plot,Year,Var,Species,Value)

plotTotal.X %>%

  # bring in plotSums
  dplyr::select(BA.Target,Plot,Year,Var,Species,Value) %>% 
  bind_rows(plotSums) %>%

  # need to convert year to numeric for plotting
  mutate(Year=as.numeric(as.character(Year))) %>%

  group_by(BA.Target,Year,Species,Var) %>% 
  summarise(Mean=mean(Value),
            SE=sd(Value)/sqrt(length(Value))) %>%

  # reorder and relabel some factors
   mutate(Var=factor(Var)) %>% 
  mutate(Var=fct_relevel(Var,"Vol","BA","sph")) %>% 
  mutate(Var=dplyr::recode(Var,BA="BA (m²/ha)",sph="SPH",
                           Vol="Vol. (m³/ha)")) %>%

  ggplot()+
  aes(x=Year,y=Mean,group=Species,fill=Species,color=Species)+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE, width=.2),
                   position=position_dodge(0))+
 scale_x_continuous(limits=c(1994,2019),breaks=c(1994,1997,2009,2019),
                     minor_breaks = c(1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis
  ylab("")+
  xlab("Measurement year")+
  facet_grid(Var~BA.Target,
             scales="free_y",
              switch="y")+
  theme(legend.position = "bottom",
  axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
  strip.placement = "outside",
            strip.background.y = element_blank())

```

`r figNums("standPlots",caption="Mean plot live-tree attributes by treatment unit and year.  Whiskers are one standard error of the mean.  Vol., volume; BA, basal area; Tree dens., tree density")`

\newpage

```{r plot periodic increment}

# Note this function performs the operation for basal area and volume only

  perInc.X %>%
  dplyr::select(BA.Target:BA.se) %>% 
    setNames(c("BA.Target","Period","Vol_mean","Vol_SE","BA_mean","BA_SE")) %>% 
    pivot_longer(cols=-c(BA.Target,Period),
                 names_to=c("Var","SE"),
                 names_sep="_",
                 values_to="Value") %>% 
    pivot_wider(names_from="SE",values_from="Value") %>%

  ggplot(aes(x=Period,y=mean,group=BA.Target,color=BA.Target))+
  geom_point()+
  geom_line()+

     geom_errorbar(aes(ymin=mean-SE, ymax=mean+SE, width=.2),
                   position=position_dodge(0))+
    facet_wrap(~Var,nrow=2,
               scales="free_y", strip.position = "left",
               labeller=as_labeller(c(BA="Basal area inc. (m²/ha/year)",Vol="Volume inc. (m³/ha/year)")))+
    ylab(NULL)+
     theme(strip.background = element_blank(),
           strip.placement = "outside")

```

`r figNums("perInc.plot",caption="Mean plot annual increment over four measurement periods for basal area and volume, by treatment.  Whiskers are one standard error of the mean. Inc, increment.")`

\newpage

```{r plot species composition by attribute}

plotTotal.table %>%
  mutate(Var=fct_relevel(Var,"BA","Vol","sph")) %>% 
  mutate(Var=dplyr::recode(Var,BA="Basal area",
                    Vol="Volume",
                    sph="Tree density")) %>%

  ggplot()+
  aes(x=Year,y=Prop,fill=Species)+
  ylab("Proportion")+
   geom_bar(position="fill",stat="identity")+
   geom_hline(yintercept=0.5, linetype="dashed", color = "black")+
  theme(legend.position = "bottom",
        axis.text.x=element_text(angle=90,vjust=0.3))+
  facet_grid(Var~BA.Target)

```

`r figNums("sppComp.plot",caption="Mean proportion of spruce and fir basal area, volume, and tree density by treatment and year.")`

\newpage

```{r plot species composition by diameter class}

# Create plotting function

diamClass.plot<-function(standVar,y.label) {
diamClass.spp %>% 
   ggplot()+
  geom_bar(aes(x=DBH.class,y=get(standVar),group=Species,fill=Species),stat="identity")+
  # geom_point(data=filter(diamClass.spp,Species=="Total"),
  #              aes(x=DBH.class,y=get(standVar)))+
  #   geom_line(data=filter(diamClass.spp,Species=="Total"),
  #              aes(x=DBH.class,y=get(standVar),group=Species))+  
   scale_fill_manual(values=c("#F8766D", "#00BFCA","azure4")) +  # control color
     theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
        legend.position = "bottom")+
    ylab(y.label)+
    xlab("DBH class (cm)")+

    # coord_flip()+
    facet_grid(Year~BA.Target) %>% 
    return()
}

# create stand-level plots by diameter distribution
diamVol.plot<-diamClass.plot("Vol",y.label="Mean plot volume (m³/ha)")
diamBA.plot<-diamClass.plot("BA",y.label="Mean plot basal area (m²/ha)")  
diamK.plot<-diamClass.plot("SPH",y.label="Mean plot tree density (sph)")

# diamClass.plot<-cowplot::plot_grid(diamVol.plot,diamBA.plot,diamK.plot,ncol=1)
```

\newpage

```{r show volume by diameter class,fig.width=8}
diamVol.plot

```

`r figNums("diamVol.plot",caption="Diameter-class distribution of volume (m³/hectare) by species, treatment and year.Numbers identifying each diameter classes represent the lower size limit for that class. For example, trees in the 7.5cm DBH class are greater or equal to 7.5 and less than 12.5cm DBH.")`

*Note that in this figure, we show volume for all diameter classes. This
is different than the stand-level estimate of volume, where we only
estimate volume for trees that would be considered merchantable under
conventional harvest in British Columbia (i.e., minimum 18.5cm DBH).*

\newpage

```{r show tree density by diameter class,fig.width=8}
diamK.plot

```

`r figNums("diamK.plot",caption="Diameter-class distribution of tree density (stems/hectare) by species, treatment and year.")`

\newpage

```{r species composition plots by diameter class}

sppComp.plot<-function(standVar){

    diamClass.spp %>% 
    filter(Species %in% c("Fir","Spruce")) %>% 
    dplyr::select(BA.Target,Species,Year,DBH.class,standVar=standVar) %>% # select columns
    pivot_wider(names_from="Species",values_from="standVar") %>%
    mutate(Fir=replace_na(Fir,0),
           Spruce=replace_na(Spruce,0)) %>%

    mutate(Total=Fir+Spruce) %>% 
    mutate(Fir.prop=Fir/Total) %>% 
    mutate(Spruce.prop=Spruce/Total) %>%

    mutate(Fir.prop=replace_na(Fir.prop,0),
           Spruce.prop=replace_na(Spruce.prop,0)) %>% 
    mutate(Spruce.prop=-Spruce.prop) %>% # change to neg value for plotting
    dplyr::select(BA.Target,Year,DBH.class,Spruce.prop,Fir.prop) %>% 
    pivot_longer(cols=contains(".prop"),names_to="Species",values_to="Prop") %>% 
    mutate(Species=str_remove(Species,".prop")) %>% 
    filter(Prop!=0) %>%

    ggplot()+
    aes(x=DBH.class,y=Prop,group=Species,fill=Species)+
      geom_bar(stat="identity")+
    xlab("DBH class (cm)")+
    scale_y_continuous(breaks=c(-1,-0.5,0,0.5,1),
                       labels=c(1,0.5,0,0.5,1))+
    ylab("Proportion")+
    coord_flip()+
    theme(legend.position="bottom")+
    facet_grid(Year~BA.Target) %>% 
    return()
} # end function

# create stand-level plots by diameter distribution
sppComp.plot("BA")

```

`r figNums("sppCompDiam.plot",caption="Spruce-fir volume proportional composition by diameter class, treatment and year.")`

*Note this figure is almost identical for compairsons of basal area,
volume or tree density. For brevity, we only show basal area.*

\newpage

```{r plot mortality as a function of 1994 stand attributes}

mortPlot.text<-paste("y = ",round(tidy(deadPlot.ANOVA)[1,"estimate"],4)," + ",round(tidy(deadPlot.ANOVA)[2,"estimate"],4),"x ",
              "\np = ",round(tidy(deadPlot.ANOVA)[2,"p.value"],4),
              "\nR² = ",round(glance(deadPlot.ANOVA)$r.squared, 3),sep="")

deadPlot.X %>% 
  ggplot()+
  aes(x=BA,y=Num.dead,color=BA.Target,shape=BA.Target)+
  geom_point(size=4)+
  stat_smooth(method = lm, formula = y ~ x,se=FALSE,aes(group=1))+
  ylab("Number of trees that died (1994-2019)")+
  xlab("Plot 1994 basal area (m²/hectare) ")+
  theme(legend.position = "bottom")+
  geom_text(aes(x=7,y=10,label=mortPlot.text,hjust=0),color="black",check_overlap = TRUE,size=4)

```

`r figNums("mort.plot",caption="Total tree mortality over 1994-2019 period vs plot basal area in 1994.  Plots are symbolized to show three treatment types.")`

\newpage

```{r plot mortality by year and DBH class}

dead %>% 
  pivot_longer(cols=BA:k,names_to = "Var",values_to = "Value") %>% 
  mutate(Var=factor(Var)) %>% 
  mutate(Var=fct_relevel(Var,"k","BA")) %>% 
  mutate(Var=dplyr::recode(Var,BA="Basal area (m²)",k="Number of trees")) %>% 
  ggplot()+
  aes(x=DBH.class,y=Value,fill=Year)+
  geom_bar(stat="identity")+
  facet_grid(Var~BA.Target,
               scales="free_y",
               switch="y")+
    ylab(NULL)+
  xlab("DBH class (cm)")+
     theme( strip.placement = "outside",
            strip.background.y = element_blank(),
            legend.position = "bottom",
            axis.text.x = element_text(angle=90,vjust = 0.3))

```

`r figNums("mortDiam.plot",caption="Total tree mortality recorded in plots from 1997-2019, presented as number of trees (top panel) and basal area (bottom panel).  Numbers identifying each diameter classes represent the lower size limit for that class. For example, trees in the 7.5cm DBH class are greater or equal to 7.5 and less than 12.5cm DBH")`

\newpage

```{r plot mortality by species and diameter class}

  deadList %>% 
  filter(Species %in% c("Spruce","Fir")) %>% 
  group_by(Species,DBH.class,BA.Target) %>% 
  summarise(Num.Dead=n()) %>%

  # plot
  ggplot()+
  aes(x=DBH.class,y=Num.Dead,fill=BA.Target) %>% 
  geom_bar(stat="identity")+
  facet_wrap(~Species)+
  ylab("Number dead 1994-2019")+
  xlab("DBH class (cm)")+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90,vjust=0.3))
```

`r figNums("mortSppDiam.plot",caption="Fir and spruce mortality recorded in plots from 1997-2019, presented as number of trees, by diameter breast height class.  Numbers identifying each diameter classes represent the lower size limit for that class.  For example, trees in the 7.5cm DBH class are greater or equal to 7.5 and less than 12.5cm DBH.")`

\newpage

# Appendices

```{r show paiTable, ft.align="left",eval=FALSE}

  tibble(`Measurement Year`=c(1992,1994,1997,2009,2019),`Growing Seasons`=c(NA,3,3,11,11)) %>% 
  flextable() %>% 
  align(align="left",part="all") %>% 
  # set_caption(caption="Number of growing seasons between measurement periods") %>% 
  autofit()

appNums("paiTable",caption="Number of growing seasons between measurement periods")
```

`r appNums("sppContrast.table",caption="Contrasts between estimated marginal means of fir and spruce growth across treatment units and initial diameter classes.  Ratio between predicted fir and growth is in 'ratio' column.")`

```{r treelevel species contrast table}

# Tree-level species contrast table
  sppContrast %>%

  mutate_if(is.double,round,4) %>%  # round to four digits
  mutate_at(vars("df","t.ratio","ratio","SE"),round,2) %>%

  # format table
  flextable(col_keys = c(colnames(.)[c(1:2,6:8,3:5)])) %>% 
  align(align="left",part = "all") %>% 
  vline( border = fp_border(color="#2c3a51" , width = 1), part = "all" ) %>% 
  hline( border = fp_border(color="#2c3a51" , width = 1), part = "all" ) %>% 
  merge_v(j=c(1:2,6:8)) %>% 
  valign(valign="top") %>%
 border_outer(part="all", border = fp_border(color="#2c3a51" , width = 2)) %>% 
   fix_border_issues()  # border(border.bottom = fp_border(color="black"))

```

\newpage

`r appNums("trtContrast.table",caption="Contrasts of predicted 1994-2019 tree growth among treatments within each species.  Ratio of predicted growth between treatments is in 'ratio' column.")`

```{r tree level treatment contrasts}

# Treatment contrast

trtContrast %>%

 mutate_if(is.double,round,4) %>%  # round to four digits
  mutate_at(vars("df","t.ratio","ratio","SE"),round,2) %>%

  # flextable
  flextable(col_keys = c(colnames(.)[c(2,1,6:8,3:5)])) %>% 
  border_outer(part="all", border = fp_border(color="#2c3a51" , width = 2)) %>% 
  align(align="left",part = "all") %>% 
  vline( border = fp_border(color="#2c3a51" , width = 1), part = "all" ) %>% 
  hline( border = fp_border(color="#2c3a51" , width = 1), part = "all" ) %>% 
  merge_v(j=c(1:2,6:8)) %>% 
  valign(valign="top") %>%
  hline_bottom() %>% 
  fix_border_issues() # border(border.bottom = fp_border(color="black"))

```

\newpage

```{r height to diameter plot}

  X %>% 
    filter(Species %in% c("Fir","Spruce")) %>% # filter for Bl and Sx for now
    filter(Status=="Live") %>% # filter for live trees
    filter(is.na(DBT)) %>% # filter for trees with no broken tops
    drop_na(Height) %>% # remove trees where height was not measured

    # Plotting
  ggplot()+
  aes(x=DBH,y=Height,fill=Species)+
  geom_point()+
  facet_grid(Year~Species)+
  theme(legend.position = "none")+
  ylab("Height (m)")+
  xlab("DBH (cm)")+
  geom_smooth(formula = y~log(x),method="lm",se=FALSE)

```

`r appNums("heightDiam.plot",caption="Relationship between height (m) and diameter (cm) for trees with measured heights, by species and year.")`

\newpage

```{r height model fit}

# Height model fit plot

  X %>% 
  filter(is.na(DBT)) %>% # filter for trees with no broken tops
  mutate(h2=exp(fixef(htDiamFit)[1]+fixef(htDiamFit)[2]*log(DBH))) %>% 
  ggplot(aes(x=Height,y=h2))+
  geom_point()+
  xlim(0,40)+
  ylim(0,40)+
  geom_smooth(method="lm",se=F)+
  ylab("Predicted height (m) from diameter-height model")+
  xlab("Measured height (m)")

```

`r appNums("htModelFit.plot",caption="Scatterplot of predicted height(m) vs actual height(m) from diameter-height model.")`

\newpage

`r appNums("standPlot.table",caption="Plot means of volume, basal area and stand density, by species, year and treatment.")`

```{r table of plot means}

standData.X %>% 
  ungroup() %>% 
  mutate_at(vars(Vol:last_col()),round,1) %>% 
  mutate_at(vars(contains("SPH")),round,0) %>%

  # change to character to allow pasting
  mutate_all(as.character) %>% 
  mutate(Vol=paste0(Vol," (",Vol.se,")")) %>% 
  mutate(BA=paste0(BA," (",BA.se,")")) %>% 
  mutate(SPH=paste0(SPH," (",SPH.se,")")) %>% 
  mutate(QMD=paste0(QMD," (",QMD.se,")")) %>%

  # rename volume column
  rename(Volume="Vol") %>%

  # remove SE columns
  dplyr::select(-contains(".se")) %>%

  # rearrange
  arrange(Year,BA.Target) %>%

  # format table
  flextable() %>%
  align(align="left",part="all") %>% 
  merge_v("Year") %>% 
  valign(valign="top",part="body") %>% 
  hline(i=seq(from=3,to=15,by=3),border=fp_border(color="black")) %>% 
  fix_border_issues() %>% 
   autofit()

```

*Volume is m3/hectare, BA = basal area (m2/ha), SPH = stems per hectare,
QMD = quadratic mean diameter (cm) at 1.3m height. Values in brackets
are the standard error of the mean.*

\newpage

`r appNums("plotPAI.table",caption="Plot means of mean annual volume, basal area and stand density increment, by species, period and treatment.")`

```{r table of periodic mean annual increment}

perInc.X %>% 
    ungroup() %>% 
  mutate_at(vars(Vol:last_col()),round,1) %>% 
  mutate_at(vars(contains("SPH")),round,0) %>%

  # change to character to allow pasting
  mutate_all(as.character) %>% 
  mutate(Vol=paste0(Vol," (",Vol.se,")")) %>% 
  mutate(BA=paste0(BA," (",BA.se,")")) %>% 
  mutate(SPH=paste0(SPH," (",SPH.se,")")) %>% 
  mutate(QMD=paste0(QMD," (",QMD.se,")")) %>%

  # rename volume column
  rename(Volume="Vol") %>%

  # remove SE columns
  dplyr::select(-contains(".se")) %>%

  # remove QMD column because it's somewhat nonsensical
  dplyr::select(-QMD) %>%

  # rearrange
  arrange(Period,BA.Target) %>%

    # Insert number of growing seasons column
  add_column(Num.Seasons=c(rep(3,3),rep(11,6)),.after="Period") %>% 
  mutate(Period=paste0(Period," (",Num.Seasons,")")) %>% 
  dplyr::select(-Num.Seasons) %>%

    # format table
  flextable() %>%
  align(align="left",part="all") %>% 
  merge_v("Period") %>% 
  valign(valign="top",part="body") %>% 
  hline(i=seq(from=3,to=9,by=3),border=fp_border(color="black")) %>% 
  fix_border_issues() %>% 
   autofit()

```

*Volume is m3/hectare, BA = basal area (m2/ha), SPH = stems per hectare.
Values in brackets in Period column are number of growing seasons in
that period. Values in brackets in Volume, BA and SPH columns are the
standard error of the mean.*

\newpage

`r appNums("standFitM.contrasts",caption="Contrasts of stand attributes (volume, basal area and tree density) between treatments and by year.  Contrasts are presented for five measurement years. SE, standard error, df, degrees freedom")`

```{r table contrasts between plot means}

standFitM.contrasts %>% 
  arrange(Contrast,Variable,Year) %>% 
  mutate_at(vars(estimate:`t.ratio`),round,2) %>% 
  mutate(p.value=round(p.value,5)) %>%

  flextable() %>% 
   align(align="left",part="all") %>% 
  merge_v(j=c("Contrast","Variable","Year")) %>% 
  valign(valign="top",part="body") %>% 
    hline(i=seq(from=5,to=40,by=5),border=fp_border(color="black")) %>% 
    fix_border_issues() %>% 
  autofit()

```

\newpage

`r appNums("standFit.contrasts",caption="Contrasts of 1992-2019 stand structure development between treatments.  Contrasts are shown for three variables: volume, basal area and stand density.")`

```{r show contrasts between plot means}

standFit.contrasts %>% 
  dplyr::select(-`Year contrast`) %>% 
  flextable() %>% 
   align(align="left",part="all") %>% 
  merge_v(j=c("Variable")) %>% 
  valign(valign="top",part="body") %>% 
    hline(i=c(3,6,9),border=fp_border(color="black")) %>% 
    fix_border_issues() %>% 
  autofit()

```

\newpage

# References